### 2020-07-10 submission/manuscript.tex
Metagenomic part in M&M done.

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
  12pt,
]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={Selective DNA and Protein Isolation from Marine Macrophyte Surfaces},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage[margin=1.0in]{geometry}
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\usepackage{times} % Times New Roman font
\usepackage[T1]{fontenc}

\usepackage[none]{hyphenat}

\usepackage{setspace}
\doublespacing
\setlength{\parskip}{1em}

\usepackage{lineno}

\usepackage{pdfpages}

\usepackage{indentfirst}

\usepackage[labelsep=period, labelfont=bf]{caption}
\renewcommand{\thefigure}{\arabic{figure}}
\renewcommand{\figurename}{Fig.}
\captionsetup{justification=raggedright,singlelinecheck=false}

\usepackage{pdflscape}
\newcommand{\blandscape}{\begin{landscape}}
\newcommand{\elandscape}{\end{landscape}}

\usepackage{siunitx}
\DeclareSIUnit\molar{\mole\per\cubic\deci\metre}
\DeclareSIUnit\Molar{\textsc{m}}
\DeclareSIUnit\cells{\text{cells}}

\usepackage{caption}
\captionsetup{justification=justified}

\usepackage{float}

\usepackage{xr}
\externaldocument[supp-]{supplementary}

\usepackage{txfonts}

\renewcommand{\figureautorefname}{Fig.}

\title{\textbf{Selective DNA and Protein Isolation from Marine Macrophyte
Surfaces}}
\author{}
\date{\vspace{-2.5em}}

\begin{document}
\maketitle

\vspace{20mm}

Marino Korlević\textsuperscript{1\(*\)}, Marsej
Markovski\textsuperscript{1}, Zihao Zhao\textsuperscript{2}, Gerhard J.
Herndl\textsuperscript{2,3}, Mirjana Najdek\textsuperscript{1}

1. Center for Marine Research, Ruđer Bošković Institute, Croatia

2. Department of Functional and Evolutionary Ecology, University of
Vienna, Austria

3. Department of Marine Microbiology and Biogeochemistry, Royal
Netherlands Institute for Sea Research, Utrecht University, The
Netherlands

\textsuperscript{\(*\)}To whom correspondence should be addressed:

Marino Korlević

G. Paliaga 5, 52210 Rovinj, Croatia

Tel.: +385 52 804 768

Fax: +385 52 804 780

e-mail:
\href{mailto:marino.korlevic@irb.hr}{\nolinkurl{marino.korlevic@irb.hr}}

Running title: DNA and protein isolation from macrophyte surfaces

\newpage
\linenumbers
\sisetup{mode=text}
\setlength\parindent{24pt}

\hypertarget{summary}{%
\subsection{Summary}\label{summary}}

\newpage

\hypertarget{introduction}{%
\subsection{Introduction}\label{introduction}}

Surfaces of marine macrophytes are inhabited by a diverse microbial
community whose structure and function are poorly understood (Egan
\emph{et al.}, 2013). As less than 1 \% of all prokaryotic species are
culturable, to study these organisms, molecular methods such as 16S rRNA
sequencing, metagenomics and metaproteomics are indispensable (Amann
\emph{et al.}, 1995; Su \emph{et al.}, 2012). Applying these techniques
requires an initial isolation step, with the purpose of obtaining high
quality DNA and proteins.

Biological material (i.e.~proteins and DNA) from pelagic microbial
communities is usually isolated by collecting cells onto filters and
subsequently isolating the aimed compound (Gilbert \emph{et al.}, 2009).
If a microbial size fraction is aimed sequential filtration is applied
(Massana \emph{et al.}, 1997; Andersson \emph{et al.}, 2010). In
contrast, obtaining biological materials from microorganism inhabiting
surfaces requires either a cell detachment procedure prior to isolation
or the host material is coextracted together with the targeted material.
Methods for separating microbial cells form the host include shaking of
host tissue (Gross \emph{et al.}, 2003; Nõges \emph{et al.}, 2010),
scraping of macrophyte surfaces (Uku \emph{et al.}, 2007) or the
application of ultrasonication (Weidner \emph{et al.}, 1996; Cai
\emph{et al.}, 2014). It was shown that shaking alone is not sufficient
to remove microbial cells, at least from plant root surfaces
(Richter-Heitmann \emph{et al.}, 2016). Manual separation methods, such
as scraping and brushing, are time consuming and subjective, as the
detachment efficiency depends on host tissue and the person performing
the procedure (Cai \emph{et al.}, 2014). Ultrasonication was proposed as
an alternative method as it is providing better results in terms of
detachment efficiency (Cai \emph{et al.}, 2014; Richter-Heitmann
\emph{et al.}, 2016). The downside of this procedure is that complete
cell removal was still not obtained and tissue disruption was observed
especially after the application of probe ultrasonication
(Richter-Heitmann \emph{et al.}, 2016). An alternative to these cell
detachment procedures is the isolation of targeted epiphytic compounds
together with host materials (Staufenberger \emph{et al.}, 2008; Jiang
\emph{et al.}, 2015). This procedure can lead to problems in the
following processing steps such as mitochondrial and chloroplast 16S
rRNA sequence contaminations from the host (Longford \emph{et al.},
2007; Staufenberger \emph{et al.}, 2008). In addition, when performing
metagenomics and metaproteomics host material can cause biased results
towards more abundant host DNA and proteins.

An alternative to these procedures is a direct isolation of the targeted
material by incubating macrophyte tissues in an extraction buffer. After
the incubation is done the undisrupted host tissue is removed and the
isolation procedure continues omitting host material contaminations. To
our knowledge the only procedure describing a direct and selective
epiphytic DNA isolation from the surfaces of marine macrophytes was
provided by Burke \emph{et al.} (2009). In contrast to previously
described methods this protocol enables an almost complete removal of
the surface community and was used for 16S rRNA gene clone libraries
construction (Burke and Thomas \emph{et al.}, 2011) and metagenomes
sequencing (Burke and Peter Steinberg \emph{et al.}, 2011). This method,
thought providing a selective isolation procedure, is using in the
extraction buffer a rapid multienzyme cleaner (3M) that is not worldwide
available and whose composition is not know (Burke \emph{et al.}, 2009).
Also to our knowledge, no selective isolation protocol for proteins from
epiphytic communities inhabiting marine macrophytes was established.

In the present study, we adapted a protocol widely used for DNA
isolation from filters (Massana \emph{et al.}, 1997) and a protocol used
for protein isolation from sediments (Chourey \emph{et al.}, 2010;
Hultman \emph{et al.}, 2015) for selective extractions of DNA and
proteins from epiphytic communities inhabiting the surfaces of two
marine macrophytes: the seagrass \emph{Cymodocea nodosa} and the alga
\emph{Caulerpa cylindracea}. In addition, we tested the removal
efficiency of the protocol and the suitability of obtained DNA and
proteins for 16S rRNA sequencing and metaproteomics.

\newpage

\hypertarget{results}{%
\subsection{Results}\label{results}}

To assess the removal efficiency of DNA and protein isolation procedures
leaves and thalli were stained with SYBR Green I and examined under a
confocal microscope before and after treatments were performed. As it is
known that marine macrophytes are harboring more algal epiphytes during
Autumn and Winter months (Reyes and Sansón, 2001), the sampling was
performed in two contrasting seasons, in December 2017 when higher
abundances of algal epiphytes are expected, and in June 2018, when lower
abundances could be expected. In addition, to further test the isolation
protocols on different macrophyte species, DNA and proteins were
isolated from the surfaces of two macrophytes, the seagrass \emph{C.
nodosa} and the macroalga \emph{C. cylindracea}. \emph{C. nodosa} was
sampled in a monospecific meadow (Bay of Saline) and \emph{C.
cylindracea} invaded meadow (Bay of Funtana), while samples of \emph{C.
cylindracea} originated from a monospecific settlement and a \emph{C.
cylindracea} invaded \emph{C. nodosa} meadow in the Bay of Funtana.
\emph{C. nodosa} June samples from the Bay of Saline were missing as the
meadow went to a decline from Spring 2018 (M. Najdek, personal
communication). Procedures developed for DNA and protein isolation
showed an almost complete removal of the surface community on both,
samples collected in December 2017 (\autoref{confocal_december}) and
June 2018 (\autoref{confocal_june}). In addition, a similar removal
efficiency was observed for communities associated with \emph{C. nodosa}
and \emph{C. cylindracea}. Also, no effect of station, settlement or
isolation procedure (DNA or protein) on the removal efficiency was
observed (Figs. \ref{confocal_december} and \ref{confocal_june}).

To evaluate if the obtained DNA is suitable to determine the microbial
community structure an Illumina sequencing of the V4 16S rRNA region was
performed. Sequencing yielded a total of 336,937 sequences after quality
curation and exclusion of eukaryotic, mitochondrial and no relative
sequences. The number of sequences classified as chloroplast was 97,328
(28.9 \si{\percent})˛ After the exclusion of these sequences the total
number of retrieved reads dropped to 239,609, ranging from 22,587 and
52,958 sequences per sample (\autoref{supp-nseq_notus}). Even when the
highest sequencing effort was applied the rarefaction curves did not
level off that is commonly observed in high-throughput 16S rRNA amplicon
sequencing procedures (\autoref{supp-rarefaction}). Sequence clustering
at a similarity level of 97 \si{\percent} yielded a total of 8,360
different OTUs. Taxonomic classification of reads allowed the
determination of the macrophyte associated epiphytic community that was
mainly composed of: \emph{Alphaproteobacteria} (14.9 ± 3.5
\si{\percent}), \emph{Bacteroidota} (12.5 ± 2.4 \si{\percent}),
\emph{Gammaproteobacteria} (11.6 ± 4.3 \si{\percent}),
\emph{Desulfobacterota} (7.8 ± 7.4 \si{\percent}), \emph{Cyanobacteria}
(6.5 ± 4.7 \si{\percent}) and \emph{Planctomycetota} (2.9 ± 1.7
\si{\percent}) (\autoref{community}).

Primers used in this study, as in many other 16S rRNA amplicon
procedures, also amplify chloroplast 16S rRNA genes. We observed a high
proportion of chloroplast sequences in all analyzed samples (33.4 ± 9.4
\si{\percent}) (\autoref{community}). To determine if chloroplast
sequences originate from hosts or eukaryotic epiphytic organisms, we
exported SILVA classified chloroplast sequences and reclassified them
using the RDP (Ribosomal Database Project) training set that allows for
a more detailed chloroplast classification. The largest proportion of
sequences was classified as Bacillariophyta (89.7 ± 5.7 \si{\percent})
indicating that the DNA removal procedure did not coextract larger
quantities of host DNA (\autoref{chloroplast}).

\hypertarget{discussion}{%
\subsection{Discussion}\label{discussion}}

\newpage

\hypertarget{experimental-procedures}{%
\subsection{Experimental procedures}\label{experimental-procedures}}

\hypertarget{sampling}{%
\subsubsection{Sampling}\label{sampling}}

Leaves of \emph{C. nodosa} were sampled in a \emph{C. nodosa} meadow in
the Bay of Saline (\ang{45;7;5} N, \ang{13;37;20} E) and in a \emph{C.
nodosa} meadow invaded by \emph{C. cylindracea} in the proximity of the
village of Funtana (\ang{45;10;39} N, \ang{13;35;42} E). Thalli of
\emph{C. cylindracea} were sampled in the same \emph{C. nodosa} invaded
meadow in Funtana and on a locality of only \emph{C. cylindracea}
located in the proximity of the invaded meadow. Leaves and thalli were
collected on the same day in two contrasting seasons, on 4 December 2017
and 18 June 2018. During spring 2018 the \emph{C. nodosa} meadow in the
Bay of Saline decayed to an extent that no leaves could be retrieved
(Najdek \emph{et al.}, unpublished data). Leaves and thalli were
collected by diving and transported to the laboratory in containers
placed on ice and filled with site seawater. Upon arrival to the
laboratory, \emph{C. nodosa} leaves were cut into sections of 1 -- 2
\si{\cm}, while \emph{C. cylindracea} thalli were cut into 5 -- 8
\si{\cm} long sections. Leaves and thalli were washed three times with
sterile artificial seawater (ASW) to remove loosely attached microbial
cells.

\hypertarget{dna-isolation}{%
\subsubsection{DNA isolation}\label{dna-isolation}}

The DNA was isolated according to the protocol for isolation from
filters described in Massana \emph{et al.} (1997). This protocol was
modified and adapted for DNA isolation from microbial communities from
macrophyte surfaces as described below. 5 \si{\ml} of lysis buffer (40
\si{\milli\Molar} EDTA, 50 \si{\milli\Molar} Tris-HCl, 0.75 \si{\Molar}
sucrose; pH 8.3) was added to 1 \si{\g} wet weight of leaves or \si{\g}
wet-weight of thalli. Lysozyme was added (final concentration 1
\si{\mg\per\ml}) and the mixture was incubated at 37 \si{\degreeCelsius}
for 30 \si{\minute}. Subsequently, proteinase K (final concentration 0.5
\si{\mg\per\ml}) and SDS (final concentration 1 \si{\percent}) were
added and the samples were incubated at 55 \si{\degreeCelsius} for 2
\si{\hour}. Following the incubation, tubes were vortexed for 10
\si{\minute} and the mixture containing lyzed epiphytic cells was
separated from host leaves or thalli by transferring the solution into a
clean tube. The lysate was extracted twice with a mixture of
phenol:chloroform:isoamyl alcohol (25:24:1; pH 8) and once with
chloroform:isoamyl alcohol (24:1). After each organic solvent mixture
addition tubes were slightly vortexed and centrifuged at 4,500 × g for
10 \si{\minute}. Following each centrifugation aqueous phases were
retrieved. After the final extraction 1/10 of chilled 3 \si{\Molar}
sodium acetate (pH 5.2) was added. DNA was precipitated by adding 1
volume of chilled isopropanol, incubating the mixtures overnight at
\num{-20} °C and centrifuging at 16,000 × g and 4 \si{\degreeCelsius}
for 20 \si{\minute}. The pellet was washed twice with 1 \si{\ml} of
chilled 70 \si{\percent} ethanol and centrifuged after each washing step
at 20,000 × g and 4 \si{\degreeCelsius} for 10 \si{\minute}. After the
first washing step duplicate pellets form the same sample were pooled
and transferred to a clean 1.5 \si{\ml} tube. The dried pellet was
resuspended in 100 \si{\ul} of deionized water.

\hypertarget{illumina-16s-rrna-sequencing}{%
\subsubsection{Illumina 16S rRNA
sequencing}\label{illumina-16s-rrna-sequencing}}

An aliquot of isolated DNA was treated with RNase A (final concentration
200 \si{\ug\per\ml}) for 2 \si{\hour} at 37 \si{\degreeCelsius}. The DNA
concentration was determined using the Quant-iT PicoGreen dsDNA Assay
Kit (Invitrogen, USA) according to the manufacturer's instructions and
diluted to 1 \si{\ng\per\ul}. The V4 region of the 16S rRNA gene was
amplified using a two-step PCR procedure. In the first PCR the 515F
(5'-GTGYCAGCMGCCGCGGTAA-3') and 806R (5'-GGACTACNVGGGTWTCTAAT-3')
primers from the Earth Microbiome Project
(\url{http://press.igsb.anl.gov/earthmicrobiome/protocols-and-standards/16s/})
were used to amplify the target region (Caporaso \emph{et al.}, 2012;
Apprill \emph{et al.}, 2015; Parada \emph{et al.}, 2016). These primers
contained on their 5' end a tagged sequence. Each sample was amplified
in four parallel 25 \si{\ul} reactions of which each contained: 1 × Q5
Reaction Buffer , 0.2 \si{\milli\Molar} of dNTPmix, 0.7 \si{\mg\per\ml}
BSA (Bovine Serum Albumin), 0.2 \si{\micro\Molar} of forward and reverse
primers, 0.5 U of Q5 High-Fidelity DNA Polymerase (New England Biolabs,
USA) and 5 \si{\ng} of DNA template. Cycling conditions were: initial
denaturation at 94 \si{\degreeCelsius} for 3 \si{\minute}, 20 cycles of
denaturation at 94 \si{\degreeCelsius} for 45 \si{\s}, annealing at 50
\si{\degreeCelsius} for 60 \si{\s} and elongation at 72
\si{\degreeCelsius} for 90 \si{\s}, finalized by an elongation step at
72 \si{\degreeCelsius} for 10 \si{\minute}. The four parallel reactions
volumes were pooled and PCR products were purified using the GeneJET PCR
Purification Kit (Thermo Fisher Scientific, USA) according to the
manufacturer's instructions and following the protocol that included
isopropanol addition for better small DNA fragment yield. The column was
eluted in 30 \si{\ul} of deionized water. Purified PCR products were
sent for Illumina MiSeq sequencing (2 × 250 bp) at IMGM Laboratories,
Martinsried, Germany. Before sequencing, at IMGM the second PCR
amplification of the two-step PCR procedure was performed using primers
targeting the tagged region incorporated in the first PCR. In addition,
these primers contained adapter and sample-specific index sequences. The
second PCR was carried out for 8 cycles. Beside samples, a positive and
negative control were sequenced. A negative control was comprised of
four parallel PCR reactions without DNA template, while for a positive
control a mock community composed of evenly mixed DNA material
originating from 20 bacterial strains (ATCC MSA-1002, ATCC, USA) was
used. Sequences obtained in this study have been deposited in the
European Nucleotide Archive (ENA) at EMBL-EBI under accession numbers
SAMEA6786270, SAMEA6648792 -- SAMEA6648794, SAMEA6648809 --
SAMEA6648811.

Obtained sequences were analyzed on the computer cluster Isabella
(University Computing Center, University of Zagreb) using mothur
(version 1.43.0) (Schloss \emph{et al.}, 2009) according to the MiSeq
Standard Operating Procedure (MiSeq SOP;
\url{https://mothur.org/wiki/MiSeq_SOP}) (Kozich \emph{et al.}, 2013)
and recommendations given from the Riffomonas project to enhance data
reproducibility (\url{http://www.riffomonas.org/}). For alignment and
classification of sequences the SILVA SSU Ref NR 99 database (release
138; \url{http://www.arb-silva.de}) was used (Quast \emph{et al.}, 2013;
Yilmaz \emph{et al.}, 2014). Sequences classified as chloroplasts by
SILVA were exported and reclassifed using mothur and the RDP
(\url{http://rdp.cme.msu.edu/}) training set (version 16) reference
files adapted for mothur (Cole \emph{et al.}, 2014). In comparison to
SILVA, RDP allows a more detailed classification of chloroplast
sequences. Pipeline data processing and visualization was done using R
(version 3.6.0) (R Core Team, 2019), package tidyverse (version 1.3.0)
(Wickham \emph{et al.}, 2019) and multiple other packages (Xie, 2014,
2015, 2020; Neuwirth, 2014; Xie \emph{et al.}, 2018; Y. Xie, 2019b,
2019a; Allaire \emph{et al.}, 2019; Zhu, 2019; Bengtsson, 2020). The
detailed analysis procedure including the R Markdown file for this paper
are available as a GitHub repository
(\url{https://github.com/mkorlevic/Korlevic_SelectiveRemoval_EnvironMicrobiol_2020}).
Based on the ATCC MSA-1002 mock community included in the analysis a
sequencing error rate of 0.009 \si{\percent} was determined, which is in
line with previously reported values for next-generation sequencing data
(Kozich \emph{et al.}, 2013; Schloss \emph{et al.}, 2016). In addition,
the negative control processed together with the samples yielded only 2
sequences after sequence quality curation.

\hypertarget{metagenomics}{%
\subsubsection{Metagenomics}\label{metagenomics}}

Four samples were selected and sent on dry ice to IMGM Laboratories,
Martinsried, Germany for metagenomic sequencing. DNA was purified using
AMPure XP Beads (Beckman Coulter, USA) applying a bead:DNA ratio of 1:1
(v/v), quantified with a Qubit dsDNA HS Assay Kit (Thermo Fisher
Scientific, USA) and check for integrity on a 1 \si{\percent} agarose
gel. Metagenomic sequencing libraries were generated from 300 \si{\ng}
of input DNA using a NEBNext Ultra II FS DNA Library Prep Kit for
Illumina (New England Biolabs, USA) according to the manufacturer's
instructions. Fragments were selected (500 -- 700 bp) using AMPure XP
Beads, PCR enriched for 3 -- 5 cycles and quality controlled. Libraries
generated from different DNA samples were pooled and sequenced on an
Illumina NovaSeq 6000 sequencing system (2 × 250 bp).

Obtained sequences were assembled individually using MEGAHIT (version
1.1.2) (Li \emph{et al.}, 2015) under default settings. Putative genes
were predicted from contings longer than 200 bp using Prodigal (version
2.6.3) (Hyatt \emph{et al.}, 2010) in metagenome mode (-p meta).
Abundances of predicted genes were expressed as Reads Per Kilobase
Million (RPKM) and calculated using the BWA algorithm (version 0.7.16a)
(Li and Durbin, 2009). All predicted genes were functionally annotated
using the eggNOG-mapper (Huerta-Cepas \emph{et al.}, 2017) and eggNOG
database (version 5.0) (Huerta-Cepas \emph{et al.}, 2019). The taxonomic
classification of sequences was determined using the lowest common
ancestor algorithm adapted from DIAMOND (version 0.8.36) (Buchfink
\emph{et al.}, 2015) and by searching against the non-redundant database
(NR). To determine phylogeny the top 10 \si{\percent} hits with an
e-value \textless{} 1 × 10\textsuperscript{-5} were used (-\/-top 10).

\hypertarget{protein-isolation}{%
\subsubsection{Protein isolation}\label{protein-isolation}}

Proteins were isolated according to the protocol for isolation from soil
described in Chourey \emph{et al.} (2010) and modified by Hultman
\emph{et al.} (2015). These protocols were further modified and adapted
for protein isolation from microbial communities form macrophyte
surfaces as described below. 20 \si{\ml} of protein extraction buffer (4
\si{\percent} SDS, 100 \si{\milli\Molar} Tris-HCl {[}pH 8.0{]}) was
added to 5 \si{\g} wet weight of leaves or 10 \si{\g} wet weight of
thalli. The mixture was incubated in boiling water for 5 \si{\minute},
vortexed for 10 \si{\minute} and incubated again in boiling water for 5
\si{\minute}. After a brief vortex the lysate was transferred to a clean
tube separating the host leaves or thalli from the mixture containing
lyzed epiphytic cells. Dithiothreitol (DTT; final concentration 24
\si{\milli\Molar}) was added and proteins were precipitated with chilled
100 \si{\percent} trichloroacetic acid (TCA; final concentration 20
\si{\percent}) overnight at \num{-20} °C. Precipitated proteins were
centrifuged at 10,000 × g and 4 \si{\degreeCelsius} for 40 \si{\minute}.
The obtained protein pellet was washed three times with chilled acetone.
During the first washing step the pellet was transferred to a clean 1.5
\si{\ml} tube. After each washing step samples were centrifuged at
20,000 × g and 4 \si{\degreeCelsius} for 5 \si{\minute}. Dried pellets
were stored at \num{-80} °C until further analysis.

\hypertarget{metaproteomics}{%
\subsubsection{Metaproteomics}\label{metaproteomics}}

Isolated proteins were whole trypsin digested using the FASP
(filter-aided sample preparation) Protein Digestion Kit (Expedeon, UK)
according to the manufacturer's instructions (Wiśniewski \emph{et al.},
2009) with small modifications. Before loading the solution to the
column, protein pellets were solubilized in a urea sample buffer
included in the kit amended with DTT (final concentration 100
\si{\milli\Molar}) for 45 \si{\minute} at room temperature and
centrifuged at 20,000 × g for 2 -- 5 \si{\minute} at room temperature to
remove larger particles. The first washing step after protein solution
loading was repeated twice. In addition, centrifugation steps were
prolonged if the column was clogged. Trypsin digestion was performed on
column filters at 37 \si{\degreeCelsius} overnight for 18 \si{\hour}.
The final filtrate containing digested proteins was acidified with 1
\si{\percent} (final concentration) trifluoroacetic acid, freezed at
\num{-80} \si{\degreeCelsius} for 15 \si{\minute}, lyophilized and sent
to the Vienna Metabolomics Center (University of Vienna) for
metaproteomic analysis. Peptides were resuspended in 1 \si{\percent}
(final concentration) trifluoroacetic acid, desalted using the Pierce
C18 Tips (Thermo Fisher Scientific, USA) according to the manufacturer's
instructions and sequenced on a Q Exactive Hybrid Quadrupole-Orbitrap
Mass Spectrometer (Thermo Fisher Scientific, USA). Obtained MS/MS
spectra were searched against a protein database from metagenomic
assembly published in Burke and Peter Steinberg \emph{et al.} (2011)
using SEQUEST-HT engines and validated with Percolator in Proteome
Discoverer 2.1 (Thermo Fisher Scientific, USA). The target-decoy
approach was used to reduce the probability of false peptide
identification. Results whose false discovery rate at the peptide level
was \SI{< 1}{\percent} were kept. For protein identification a minimum
of of two peptides and one unique peptide were required. For protein
quantification, a chromatographic peak area-based free quantitative
method was applied.

\hypertarget{confocal-microscopy}{%
\subsubsection{Confocal microscopy}\label{confocal-microscopy}}

Host leaves and thalli from DNA and protein isolation steps were washed
seven times in deionized water and fixed with formaldehyde (final
concentration \textasciitilde{} 3 \si{\percent}). In addition,
nontreated leaves and thalli, washed three times in ASW to remove
loosely attached microbial cells, were fixed in the same concentration
of formaldehyde and used as a positive control. For long therm storage,
fixed leaves and thalli were immersed in a mixture of phosphate-buffered
saline (PBS) and ethanol (1:1) and stored at \num{-20}
\si{\degreeCelsius}. Treated and untreated leaves and thalli segments
were stained in a 2 × solution of SYBR Green I and examined under a
Leica TCS SP8 X FLIM confocal microscope (Leica Microsystems, Germany).

\hypertarget{acknowledgements}{%
\subsection{Acknowledgements}\label{acknowledgements}}

\newpage

\hypertarget{references}{%
\subsection{References}\label{references}}

\hypertarget{refs}{}
\leavevmode\hypertarget{ref-Allaire2019}{}%
Allaire, J.J., Xie, Y., McPherson, J., Luraschi, J., Ushey, K., Atkins,
A., et al. (2019) rmarkdown: Dynamic documents for R.

\leavevmode\hypertarget{ref-Amann1995}{}%
Amann, R.I., Ludwig, W., and Schleifer, K.H. (1995) Phylogenetic
identification and \emph{in situ} detection of individual microbial
cells without cultivation. \emph{Microbiol Rev} \textbf{59}: 143--169.

\leavevmode\hypertarget{ref-Andersson2010}{}%
Andersson, A.F., Riemann, L., and Bertilsson, S. (2010) Pyrosequencing
reveals contrasting seasonal dynamics of taxa within Baltic Sea
bacterioplankton communities. \emph{ISME J} \textbf{4}: 171--181.

\leavevmode\hypertarget{ref-Apprill2015}{}%
Apprill, A., McNally, S., Parsons, R., and Weber, L. (2015) Minor
revision to V4 region SSU rRNA 806R gene primer greatly increases
detection of SAR11 bacterioplankton. \emph{Aquat Microb Ecol}
\textbf{75}: 129--137.

\leavevmode\hypertarget{ref-Bengtsson2020}{}%
Bengtsson, H. (2020) MatrixStats: Functions that apply to rows and
columns of matrices (and to vectors).

\leavevmode\hypertarget{ref-Buchfink2015}{}%
Buchfink, B., Xie, C., and Huson, D.H. (2015) Fast and sensitive protein
alignment using DIAMOND. \emph{Nat Methods} \textbf{12}: 59--60.

\leavevmode\hypertarget{ref-Burke2009}{}%
Burke, C., Kjelleberg, S., and Thomas, T. (2009) Selective extraction of
bacterial DNA from the surfaces of macroalgae. \emph{Appl Environ
Microbiol} \textbf{75}: 252--256.

\leavevmode\hypertarget{ref-Burke2011}{}%
Burke, C., Steinberg, P., Rusch, D., Kjelleberg, S., and Thomas, T.
(2011) Bacterial community assembly based on functional genes rather
than species. \emph{Proc Natl Acad Sci U S A} \textbf{108}:
14288--14293.

\leavevmode\hypertarget{ref-Burke2011a}{}%
Burke, C., Thomas, T., Lewis, M., Steinberg, P., and Kjelleberg, S.
(2011) Composition, uniqueness and variability of the epiphytic
bacterial community of the green alga \emph{Ulva australis}. \emph{ISME
J} \textbf{5}: 590--600.

\leavevmode\hypertarget{ref-Cai2014}{}%
Cai, X., Gao, G., Yang, J., Tang, X., Dai, J., Chen, D., and Song, Y.
(2014) An ultrasonic method for separation of epiphytic microbes from
freshwater submerged macrophytes. \emph{J Basic Microbiol} \textbf{54}:
758--761.

\leavevmode\hypertarget{ref-Caporaso2012}{}%
Caporaso, J.G., Lauber, C.L., Walters, W.A., Berg-Lyons, D., Huntley,
J., Fierer, N., et al. (2012) Ultra-high-throughput microbial community
analysis on the Illumina HiSeq and MiSeq platforms. \emph{ISME J}
\textbf{6}: 1621--1624.

\leavevmode\hypertarget{ref-Chourey2010}{}%
Chourey, K., Jansson, J., VerBerkmoes, N., Shah, M., Chavarria, K.L.,
Tom, L.M., et al. (2010) Direct cellular lysis/protein extraction
protocol for soil metaproteomics. \emph{J Proteome Res} \textbf{9}:
6615--6622.

\leavevmode\hypertarget{ref-Cole2014}{}%
Cole, J.R., Wang, Q., Fish, J.A., Chai, B., McGarrell, D.M., Sun, Y., et
al. (2014) Ribosomal Database Project: Data and tools for high
throughput rRNA analysis. \emph{Nucleic Acids Res} \textbf{42}:
D633--D642.

\leavevmode\hypertarget{ref-Egan2013}{}%
Egan, S., Harder, T., Burke, C., Steinberg, P., Kjelleberg, S., and
Thomas, T. (2013) The seaweed holobiont: Understanding seaweed-bacteria
interactions. \emph{FEMS Microbiol Rev} \textbf{37}: 462--476.

\leavevmode\hypertarget{ref-Gilbert2009}{}%
Gilbert, J.A., Field, D., Swift, P., Newbold, L., Oliver, A., Smyth, T.,
et al. (2009) The seasonal structure of microbial communities in the
Western English Channel. \emph{Environ Microbiol} \textbf{11}:
3132--3139.

\leavevmode\hypertarget{ref-Gross2003}{}%
Gross, E.M., Feldbaum, C., and Graf, A. (2003) Epiphyte biomass and
elemental composition on submersed macrophytes in shallow eutrophic
lakes. \emph{Hydrobiologia} \textbf{506-509}: 559--565.

\leavevmode\hypertarget{ref-Huerta-Cepas2017}{}%
Huerta-Cepas, J., Forslund, K., Coelho, L.P., Szklarczyk, D., Jensen,
L.J., von Mering, C., and Bork, P. (2017) Fast genome-wide functional
annotation through orthology assignment by eggNOG-mapper. \emph{Mol Biol
Evol} \textbf{34}: 2115--2122.

\leavevmode\hypertarget{ref-Huerta-Cepas2019}{}%
Huerta-Cepas, J., Szklarczyk, D., Heller, D., Hernández-Plaza, A.,
Forslund, S.K., Cook, H., et al. (2019) eggNOG 5.0: A hierarchical,
functionally and phylogenetically annotated orthology resource based on
5090 organisms and 2502 viruses. \emph{Nucleic Acids Res} \textbf{47}:
D309--D314.

\leavevmode\hypertarget{ref-Hultman2015}{}%
Hultman, J., Waldrop, M.P., Mackelprang, R., David, M.M., McFarland, J.,
Blazewicz, S.J., et al. (2015) Multi-omics of permafrost, active layer
and thermokarst bog soil microbiomes. \emph{Nature} \textbf{521}:
208--212.

\leavevmode\hypertarget{ref-Hyatt2010}{}%
Hyatt, D., Chen, G.-L., LoCascio, P.F., Land, M.L., Larimer, F.W., and
Hauser, L.J. (2010) Prodigal: Prokaryotic gene recognition and
translation initiation site identification. \emph{BMC Bioinformatics}
\textbf{11}: 119.

\leavevmode\hypertarget{ref-Jiang2015}{}%
Jiang, Y.-F., Ling, J., Dong, J.-D., Chen, B., Zhang, Y.-Y., Zhang,
Y.-Z., and Wang, Y.-S. (2015) Illumina-based analysis the microbial
diversity associated with \emph{Thalassia hemprichii} in Xincun Bay,
South China Sea. \emph{Ecotoxicology} \textbf{24}: 1548--56.

\leavevmode\hypertarget{ref-Kozich2013}{}%
Kozich, J.J., Westcott, S.L., Baxter, N.T., Highlander, S.K., and
Schloss, P.D. (2013) Development of a dual-index sequencing strategy and
curation pipeline for analyzing amplicon sequence data on the MiSeq
Illumina sequencing platform. \emph{Appl Environ Microbiol} \textbf{79}:
5112--5120.

\leavevmode\hypertarget{ref-Li2015}{}%
Li, D., Liu, C.-M., Luo, R., Sadakane, K., and Lam, T.-W. (2015)
MEGAHIT: An ultra-fast single-node solution for large and complex
metagenomics assembly via succinct de Bruijn graph.
\emph{Bioinformatics} \textbf{31}: 1674--1676.

\leavevmode\hypertarget{ref-Li2009a}{}%
Li, H. and Durbin, R. (2009) Fast and accurate short read alignment with
BurrowsWheeler transform. \emph{Bioinformatics} \textbf{25}: 1754--1760.

\leavevmode\hypertarget{ref-Longford2007}{}%
Longford, S., Tujula, N., Crocetti, G., Holmes, A., Holmström, C.,
Kjelleberg, S., et al. (2007) Comparisons of diversity of bacterial
communities associated with three sessile marine eukaryotes. \emph{Aquat
Microb Ecol} \textbf{48}: 217--229.

\leavevmode\hypertarget{ref-Massana1997}{}%
Massana, R., Murray, A.E., Preston, C.M., and DeLong, E.F. (1997)
Vertical distribution and phylogenetic characterization of marine
planktonic \emph{Archaea} in the Santa Barbara Channel. \emph{Appl
Environ Microbiol} \textbf{63}: 50--56.

\leavevmode\hypertarget{ref-Neuwirth2014}{}%
Neuwirth, E. (2014) RColorBrewer: ColorBrewer palettes.

\leavevmode\hypertarget{ref-Noges2010}{}%
Nõges, T., Luup, H., and Feldmann, T. (2010) Primary production of
aquatic macrophytes and their epiphytes in two shallow lakes (Peipsi and
Võrtsjärv) in Estonia. \emph{Aquat Ecol} \textbf{44}: 83--92.

\leavevmode\hypertarget{ref-Parada2016}{}%
Parada, A.E., Needham, D.M., and Fuhrman, J.A. (2016) Every base
matters: Assessing small subunit rRNA primers for marine microbiomes
with mock communities, time series and global field samples.
\emph{Environ Microbiol} \textbf{18}: 1403--1414.

\leavevmode\hypertarget{ref-Quast2013}{}%
Quast, C., Pruesse, E., Yilmaz, P., Gerken, J., Schweer, T., Yarza, P.,
et al. (2013) The SILVA ribosomal RNA gene database project: Improved
data processing and web-based tools. \emph{Nucleic Acids Res}
\textbf{41}: D590--D596.

\leavevmode\hypertarget{ref-RCoreTeam2019}{}%
R Core Team (2019) R: A language and environment for statistical
computing, Vienna, Austria: R Foundation for Statistical Computing.

\leavevmode\hypertarget{ref-Reyes2001}{}%
Reyes, J. and Sansón, M. (2001) Biomass and production of the epiphytes
on the leaves of \emph{Cymodocea nodosa} in the Canary Islands.
\emph{Bot Mar} \textbf{44}: 307--313.

\leavevmode\hypertarget{ref-Richter-Heitmann2016}{}%
Richter-Heitmann, T., Eickhorst, T., Knauth, S., Friedrich, M.W., and
Schmidt, H. (2016) Evaluation of strategies to separate root-associated
microbial communities: A crucial choice in rhizobiome research.
\emph{Front Microbiol} \textbf{7}: 773.

\leavevmode\hypertarget{ref-Schloss2016}{}%
Schloss, P.D., Jenior, M.L., Koumpouras, C.C., Westcott, S.L., and
Highlander, S.K. (2016) Sequencing 16S rRNA gene fragments using the
PacBio SMRT DNA sequencing system. \emph{PeerJ} \textbf{4}: e1869.

\leavevmode\hypertarget{ref-Schloss2009}{}%
Schloss, P.D., Westcott, S.L., Ryabin, T., Hall, J.R., Hartmann, M.,
Hollister, E.B., et al. (2009) Introducing mothur: Open-source,
platform-independent, community-supported software for describing and
comparing microbial communities. \emph{Appl Environ Microbiol}
\textbf{75}: 7537--7541.

\leavevmode\hypertarget{ref-Staufenberger2008}{}%
Staufenberger, T., Thiel, V., Wiese, J., and Imhoff, J.F. (2008)
Phylogenetic analysis of bacteria associated with \emph{Laminaria
saccharina}. \emph{FEMS Microbiol Ecol} \textbf{64}: 65--77.

\leavevmode\hypertarget{ref-Su2012}{}%
Su, C., Lei, L., Duan, Y., Zhang, K.-Q., and Yang, J. (2012)
Culture-independent methods for studying environmental microorganisms:
Methods, application, and perspective. \emph{Appl Microbiol Biotechnol}
\textbf{93}: 993--1003.

\leavevmode\hypertarget{ref-Uku2007}{}%
Uku, J., Björk, M., Bergman, B., and Díez, B. (2007) Characterization
and comparison of prokaryotic epiphytes associated with three East
African seagrasses. \emph{J Phycol} \textbf{43}: 768--779.

\leavevmode\hypertarget{ref-Weidner1996}{}%
Weidner, S., Arnold, W., and Puhler, A. (1996) Diversity of uncultured
microorganisms associated with the seagrass \emph{Halophila stipulacea}
estimated by restriction fragment length polymorphism analysis of
PCR-amplified 16S rRNA genes. \emph{Appl Environ Microbiol} \textbf{62}:
766--771.

\leavevmode\hypertarget{ref-Wickham2019}{}%
Wickham, H., Averick, M., Bryan, J., Chang, W., McGowan, L.D., François,
R., et al. (2019) Welcome to the tidyverse. \emph{J Open Source Softw}
\textbf{4}: 1686.

\leavevmode\hypertarget{ref-Wisniewski2009}{}%
Wiśniewski, J.R., Zougman, A., Nagaraj, N., and Mann, M. (2009)
Universal sample preparation method for proteome analysis. \emph{Nat
Methods} \textbf{6}: 359--362.

\leavevmode\hypertarget{ref-Xie2015}{}%
Xie, Y. (2015) Dynamic Documents with R and knitr, 2nd ed. Boca Raton,
Florida: Chapman and Hall/CRC.

\leavevmode\hypertarget{ref-Xie2014}{}%
Xie, Y. (2014) knitr: A comprehensive tool for reproducible research in
R. In \emph{Implementing Reproducible Computational Research}. Stodden,
V., Leisch, F., and Peng, R.D. (eds). New York: Chapman and Hall/CRC,
pp. 3--32.

\leavevmode\hypertarget{ref-Xie2019}{}%
Xie, Y. (2019a) knitr: A general-purpose package for dynamic report
generation in R.

\leavevmode\hypertarget{ref-Xie2019a}{}%
Xie, Y. (2019b) TinyTeX: A lightweight, cross-platform, and
easy-to-maintain LaTeX distribution based on TeX Live. \emph{TUGboat}
\textbf{40}: 30--32.

\leavevmode\hypertarget{ref-Xie2020}{}%
Xie, Y. (2020) tinytex: Helper functions to install and maintain 'TeX
Live', and compile 'LaTeX' documents.

\leavevmode\hypertarget{ref-Xie2018}{}%
Xie, Y., Allaire, J.J., and Grolemund, G. (2018) R Markdown: The
Definitive Guide, 1st ed. Boca Raton, Florida: Chapman and Hall/CRC.

\leavevmode\hypertarget{ref-Yilmaz2014}{}%
Yilmaz, P., Parfrey, L.W., Yarza, P., Gerken, J., Pruesse, E., Quast,
C., et al. (2014) The SILVA and "All-Species Living Tree Project (LTP)"
taxonomic frameworks. \emph{Nucleic Acids Res} \textbf{42}: D643--D648.

\leavevmode\hypertarget{ref-Zhu2019}{}%
Zhu, H. (2019) kableExtra: Construct complex table with 'kable' and pipe
syntax.

\newpage 
\setlength\parindent{0pt}

\hypertarget{figure-captions}{%
\subsection{Figure Captions}\label{figure-captions}}

\textbf{\autoref{confocal_december}.} \nameref{confocal_december}

\textbf{\autoref{confocal_june}.} \nameref{confocal_june}

\textbf{\autoref{community}.} \nameref{community}

\textbf{\autoref{chloroplast}.} \nameref{chloroplast}

\newpage

\hypertarget{figures}{%
\subsection{Figures}\label{figures}}

\begin{figure}[ht]

{\centering \includegraphics[width=0.85\linewidth]{../results/images/confocal_december}

}

\caption{Confocal microscope images of \textit{C. nodosa} and \textit{C. cylindracea} surfaces from the Bay of Saline and the Bay of Funtana (Mixed and Monospecific Settlements) sampled on 4 December 2017 and stained with SYBR Green I. Scale bar in all images is 60 \si{\um}.\label{confocal_december}}\label{fig:unnamed-chunk-2}
\end{figure}

\newpage
\begin{figure}[ht]

{\centering \includegraphics[width=0.85\linewidth]{../results/images/confocal_june}

}

\caption{Confocal microscope images of \textit{C. nodosa} and \textit{C. cylindracea} surfaces from the Bay of Funtana (Mixed and Monospecific Settlements) sampled on 19 June 2018 and stained with SYBR Green I. Scale bar in all images is 60 \si{\um}.\label{confocal_june}}\label{fig:unnamed-chunk-3}
\end{figure}

\newpage
\begin{figure}[ht]

{\centering \includegraphics[width=1\linewidth]{../results/figures/community_bar_plot}

}

\caption{Taxonomic classification and relative contribution of the most abundant (> 1 \si{\percent}) bacterial and archaeal sequences from surfaces of two marine macrophytes (\textit{C. nodosa} and \textit{C. cylindracea}) sampled in the Bay of Saline and the Bay of Funtana (Mixed and Monospecific Settlements) and in two contrasting seasons (4 December 2017 and 19 June 2018).\label{community}}\label{fig:unnamed-chunk-4}
\end{figure}

\newpage
\begin{figure}[ht]

{\centering \includegraphics[width=1\linewidth]{../results/figures/chloroplast}

}

\caption{Taxonomic classification and relative contribution of chloroplast sequences from surfaces of two marine macrophytes (\textit{C. nodosa} and \textit{C. cylindracea}) sampled in the Bay of Saline and the Bay of Funtana (Mixed and Monospecific Settlements) and in two contrasting seasons (4 December 2017 and 19 June 2018).\label{chloroplast}}\label{fig:unnamed-chunk-5}
\end{figure}

\end{document}