### 2020-11-16 submission/manuscript.tex

\documentclass[12pt,]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=1.0in]{geometry}
\usepackage{hyperref}
\hypersetup{unicode=true,
            pdftitle={Selective DNA and Protein Isolation from Marine Macrophyte Surfaces},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

%%% Use protect on footnotes to avoid problems with footnotes in titles
\let\rmarkdownfootnote\footnote%
\def\footnote{\protect\rmarkdownfootnote}

%%% Change title format to be more compact
\usepackage{titling}

% Create subtitle command for use in maketitle
\providecommand{\subtitle}[1]{
  \posttitle{
    \begin{center}\large#1\end{center}
    }
}

\setlength{\droptitle}{-2em}

  \title{\textbf{Selective DNA and Protein Isolation from Marine Macrophyte
Surfaces}}
    \pretitle{\vspace{\droptitle}\centering\huge}
  \posttitle{\par}
    \author{}
    \preauthor{}\postauthor{}
    \date{}
    \predate{}\postdate{}

\usepackage{times} % Times New Roman font
\usepackage[T1]{fontenc}

\usepackage[none]{hyphenat}

\usepackage{setspace}
\doublespacing
\setlength{\parskip}{1em}

\usepackage{lineno}
\renewcommand{\linenumberfont}{\normalfont\tiny}

\usepackage{pdfpages}

\usepackage{indentfirst}

\usepackage[labelsep=period, labelfont=bf]{caption}
\renewcommand{\thefigure}{\arabic{figure}}
\renewcommand{\figurename}{Fig.}
\captionsetup{justification=raggedright,singlelinecheck=false}

\usepackage{pdflscape}
\newcommand{\blandscape}{\begin{landscape}}
\newcommand{\elandscape}{\end{landscape}}

\usepackage{siunitx}
\DeclareSIUnit\molar{\mole\per\cubic\deci\metre}
\DeclareSIUnit\Molar{\textsc{m}}
\DeclareSIUnit\cells{\text{cells}}

\usepackage{caption}
\captionsetup{justification=justified}

\usepackage{float}

\usepackage{xr}
\externaldocument[supp-]{supplementary}

\usepackage{txfonts}

\renewcommand{\figureautorefname}{Fig.}

\begin{document}
\maketitle

\vspace{20mm}

Marino Korlević\textsuperscript{1\(*\)}, Marsej
Markovski\textsuperscript{1}, Zihao Zhao\textsuperscript{2}, Gerhard J.
Herndl\textsuperscript{2,3,4}, Mirjana Najdek\textsuperscript{1}

1. Center for Marine Research, Ruđer Bošković Institute, Croatia

2. Department of Functional and Evolutionary Ecology, University of
Vienna, Austria

3. NIOZ, Department of Marine Microbiology and Biogeochemistry, Royal
Netherlands Institute for Sea Research, Utrecht University, The
Netherlands

4. Vienna Metabolomics Center, University of Vienna, Austria

\textsuperscript{\(*\)}To whom correspondence should be addressed:

Marino Korlević

G. Paliaga 5, 52210 Rovinj, Croatia

Tel.: +385 52 804 768

Fax: +385 52 804 780

e-mail:
\href{mailto:marino.korlevic@irb.hr}{\nolinkurl{marino.korlevic@irb.hr}}

Running title: DNA and protein isolation from macrophyte surfaces

\newpage

\sisetup{mode=text}
\setlength\parindent{24pt}

\hypertarget{summary}{%
\subsection{Summary}\label{summary}}

Studies of unculturable microbes often combine methods based on DNA,
such as 16S rRNA sequencing and metagenomics, with methods that allow
insights into the metabolic status, such as metaproteomics. To apply
these techniques to the microbial community inhabiting the surfaces of
marine macrophytes it is advisable to perform a selective DNA and
protein isolation prior to the analysis to avoid biases due to the host
material being present in high quantities. Two protocols for DNA and
protein isolation were adapted for selective extractions of DNA and
proteins from epiphytic communities inhabiting the surfaces of two
marine macrophytes, the seagrass \emph{Cymodocea nodosa} and the
macroalga \emph{Caulerpa cylindracea}. Protocols showed an almost
complete removal of the epiphytic community regardless of the sampling
season, station, settlement or host species. The obtained DNA was
suitable for metagenomic and 16S rRNA sequencing, while isolated
proteins could be identified by mass spectrometry. The low presence of
host DNA and proteins in the samples indicated a high specificity of the
protocols. The procedures are based on universally available laboratory
chemicals making the protocols widely applicable. Taken together, the
adapted protocols ensure an almost complete removal of the macrophyte
epiphytic community. The procedures are selective for microbes
inhabiting macrophyte surfaces and provide DNA and proteins applicable
in 16S rRNA sequencing, metagenomics and metaproteomics.

\newpage

\hypertarget{introduction}{%
\subsection{Introduction}\label{introduction}}

Surfaces of marine macrophytes are colonized by a diverse microbial
community whose structure and function are poorly understood (Egan
\emph{et al.}, 2013). As less than 1 \% of all prokaryotic species are
culturable, molecular methods such as 16S rRNA sequencing, metagenomics
and metaproteomics are indispensable to study these organisms (Amann
\emph{et al.}, 1995; Su \emph{et al.}, 2012). Applying these techniques
requires an initial isolation step with the purpose of obtaining high
quality DNA and proteins.

Biological material (i.e., proteins and DNA) from pelagic microbial
communities is usually isolated by collecting cells onto filters and
subsequently isolating the target organisms or communities (Gilbert
\emph{et al.}, 2009). If a specific microbial size fraction is aimed
sequential filtration is applied (Massana \emph{et al.}, 1997; Andersson
\emph{et al.}, 2010). In contrast, obtaining microorganisms associated
to surfaces require either a cell detachment procedure prior to
isolation or the host material is co-extracted with the target material.
Methods for separating microbial cells from the host include shaking of
host tissue (Gross \emph{et al.}, 2003; Nõges \emph{et al.}, 2010),
scraping of macrophyte surfaces (Uku \emph{et al.}, 2007) or applying
ultrasonication (Weidner \emph{et al.}, 1996; Cai \emph{et al.}, 2014).
It was shown that shaking alone is not sufficient to remove microbial
cells from surfaces, at least not from plant root surfaces
(Richter-Heitmann \emph{et al.}, 2016). Manual separation methods, such
as scraping and brushing are time consuming and subjective, as the
detachment efficiency depends on host tissue and the person performing
the procedure (Cai \emph{et al.}, 2014). Ultrasonication was proposed as
an alternative method as it is providing better results in terms of
detachment efficiency (Cai \emph{et al.}, 2014; Richter-Heitmann
\emph{et al.}, 2016). The downside of this procedure is that complete
cell removal is still not obtained and tissue disruption was observed
especially after the application of probe ultrasonication
(Richter-Heitmann \emph{et al.}, 2016). An alternative to these cell
detachment procedures is the isolation of target epiphytic compounds
together with host material (Staufenberger \emph{et al.}, 2008; Jiang
\emph{et al.}, 2015). This procedure can lead to problems in the
following processing steps such as mitochondrial and chloroplast 16S
rRNA sequence contaminations from the host (Longford \emph{et al.},
2007; Staufenberger \emph{et al.}, 2008). In addition, when performing
metagenomics and metaproteomics host material can cause biased results
towards more abundant host DNA and proteins.

An alternative to these procedures is a direct isolation of the target
material by incubating macrophyte tissues in an extraction buffer. After
the incubation, the undisrupted host tissue is removed followed by the
isolation procedure, omitting host material contaminations. To our
knowledge, the only procedure describing a direct and selective
epiphytic DNA isolation from the surfaces of marine macrophytes was
described by Burke \emph{et al.} (2009). In contrast to previously
described methods, this protocol enables an almost complete removal of
the surface community. It was used for 16S rRNA gene clone library
construction (Burke and Thomas \emph{et al.}, 2011) and metagenome
sequencing (Burke and Peter Steinberg \emph{et al.}, 2011). This method,
although providing a selective isolation procedure, uses a rapid
multi-enzyme cleaner (3M) that is not available worldwide and the
chemical constituents are unknown (Burke \emph{et al.}, 2009). Also to
our knowledge, no selective isolation protocol to perform
(meta)proteomics of epiphytic communities associated with marine
macrophytes has been developed yet.

In the present study, we adapted a protocol widely used for DNA
isolation from filters (Massana \emph{et al.}, 1997) and a protocol used
for protein isolation from soils (Chourey \emph{et al.}, 2010; Hultman
\emph{et al.}, 2015). These two adapted methods allowed for a selective
extraction of DNA and proteins from epiphytic communities inhabiting the
surfaces of two marine macrophytes, the seagrass \emph{Cymodocea nodosa}
and the macroalga \emph{Caulerpa cylindracea}. In addition, we tested
the removal efficiency of the protocol and the suitability of obtained
DNA and proteins for 16S rRNA sequencing, metagenomics and
metaproteomics.

\newpage

\hypertarget{results}{%
\subsection{Results}\label{results}}

To assess the removal efficiency of the DNA and protein isolation
procedures, leaves and thalli were examined under a confocal microscope
before and after treatments were performed. The modified procedures
resulted in an almost complete removal of the surface community of both,
\emph{C. nodosa} and \emph{C. cylindracea}. In addition, a similar
removal efficiency was observed for communities sampled in contrasting
months, December 2017 (\autoref{confocal_december}) and June 2018
(\autoref{confocal_june}). Also, no effect of station, settlement or
isolation procedure (DNA or protein) on the removal efficiency was
observed (Figs. \ref{confocal_december} and \ref{confocal_june}).

To evaluate whether the obtained DNA is suitable to determine the
composition of the microbial community Illumina sequencing of the V4
region of the 16S rRNA was performed. Sequencing yielded a total of
336,944 sequences after quality curation and exclusion of eukaryotic,
mitochondrial and no relative sequences. The number of sequences
classified as chloroplasts was 97,334. After excluding these sequences
the total number of retrieved reads was 239,610, ranging from 22,596 to
52,930 sequences per sample (\autoref{supp-nseq_notus}). Even when the
highest sequencing effort was applied, the rarefaction curves did not
level off which is commonly observed in high-throughput 16S rRNA
amplicon sequencing (\autoref{supp-rarefaction}). Sequences clustering
at a similarity level of 97 \si{\percent} yielded a total of 8,355
different OTUs. Taxonomic classification of reads revealed a
macrophyte-associated epiphytic community that mainly composed of
\emph{Alphaproteobacteria} (14.9 ± 3.5 \si{\percent}),
\emph{Bacteroidota} (12.5 ± 2.4 \si{\percent}),
\emph{Gammaproteobacteria} (11.6 ± 4.3 \si{\percent}),
\emph{Desulfobacterota} (7.8 ± 7.5 \si{\percent}), \emph{Cyanobacteria}
(6.5 ± 4.7 \si{\percent}) and \emph{Planctomycetota} (2.9 ± 1.7
\si{\percent}) (\autoref{community}).

Primers used in this study, as in many other 16S rRNA amplicon
procedures, also amplified chloroplast 16S rRNA genes. We observed a
high proportion of chloroplast sequences in all analyzed samples (33.4 ±
9.4 \si{\percent}) (\autoref{community}). To determine whether
chloroplast sequences originate from the host or eukaryotic epiphytic
organisms, we exported SILVA-classified chloroplast sequences and
reclassified them using the RDP (Ribosomal Database Project) training
set that allows for a more detailed chloroplast classification. The
largest proportion of sequences was classified as Bacillariophyta (89.7
± 5.7 \si{\percent}) indicating that the DNA removal procedure resulted
in only minor co-extracted quantities of host DNA
(\autoref{chloroplast}). Chloroplast sequences classified as
Streptophyta constituted 3.3 ± 2.8 \si{\percent} of all chloroplast
sequences originating from \emph{C. nodosa} samples, while sequences
classified as Chlorophyta comprised only 0.02 ± 0.01 \si{\percent} of
all chloroplast sequences associated with \emph{C. cylindracea} samples.

To determine whether the extracted DNA can be used for metagenomic
sequencing, four samples containing epiphytic DNA were selected and
shotgun sequenced using an Illumina platform. Metagenomic sequencing
yielded between 207,149,524 and 624,029,930 sequence pairs
(\autoref{supp-metagenomic_statistics}). Obtained sequences were
successfully assembled into contigs whose L50 ranged from 654 to 1,011
bp. In addition, predicted coding sequences were functionally annotated
(9,066,667 -- 20,256,215 annotated sequences; \autoref{cog}) and
taxonomically classified. Functional annotation allowed for an
assessment of the relative contribution of each COG (Clusters of
Orthologous Groups) functional category to the total number of annotated
coding sequences (\autoref{cog}a). Functional categories containing the
highest number of sequences were C (Energy production and conversion), E
(Amino acid transport and metabolism), M (Cell wall/membrane/envelope
biogenesis), L (Replication, recombination and repair) and P (Inorganic
ion transport and metabolism). If host DNA is co-extracted with
epiphytes it should be detected in large proportions in sequenced
metagenomes. However, no large proportions of coding sequences
classified as Streptophyta and Chlorophyta were detected
(\autoref{supp-metagenomic_taxonomy}). Sequenced metagenomic DNA
originating from the surface of \emph{C. nodosa} contained 1.3
\si{\percent} of coding sequences classified as Streptophyta in December
2017 and 0.7 \si{\percent} in June 2018. Furthermore, the summed RPKM
(Reads Per Kilobase Million) of these sequences constituted 1.7
\si{\percent} of total RPKM of all successfully classified sequences in
December 2017 and 1.1 \si{\percent} in June 2018. Similar low
proportions of host coding sequences were detected in metagenomic
samples originating from the surfaces of \emph{C. cylindracea}. Of all
successfully classified coding sequences 0.2 \si{\percent} were
classified as Chlorophyta in December 2017 and 0.1 \si{\percent} in June
2018. A relatively higher proportion of RPKM of these sequences than in
the case of \emph{C. nodosa} was observed, indicating a higher
coextraction of host DNA in \emph{C. cylindracea}. In December, the
proportion of RPKM of sequences classified as Chlorophyta was 8.2
\si{\percent}, while in June 2018 it reached 13.6 \si{\percent}.

To evaluate whether the procedure for protein extraction is suitable for
metaproteomic analysis, obtained proteins were trypsin digested and
sequenced using a mass spectrometer. Obtained MS/MS spectra were
searched against a protein database from sequenced metagenomes. From
14,219 to 16,449 proteins were identified in isolated protein samples
(\autoref{cog}b). In addition, successful identification of proteins
allowed for an assessment of the relative contribution of each COG
functional category to the total number of identified proteins
(\autoref{cog}b). Functional categories containing the highest number of
identified proteins were C (Energy production and conversion), G
(Carbohydrate transport and metabolism), P (Inorganic ion transport and
metabolism), O (Posttranslational modification, protein turnover,
chaperones) and E (Amino acid transport and metabolism). Isolated
proteins could originate from epiphytic organisms inhabiting the
macrophyte surface and/or from macrophyte tissue underlying them. The
contribution of proteins originating from host tissues was evaluated by
identifying all the proteins predicted to belong to any taxonomic group
within the phyla Streptophyta and Chlorophyta and by calculating their
contribution to the number and abundance (NAAF -- Normalized Abundance
Area Factor) of all identified proteins. On average, proteins isolated
from the surface of \emph{C. nodosa} contained 1.8 ± 0.06 \si{\percent}
of proteins associated with Streptophyta, contributing to 2.2 ± 0.8
\si{\percent} of total proteins. Similar to metagenomes, proteins
associated with Chlorophyta contributed more to \emph{C. cylindracea}
than proteins associated with Streptophyta to \emph{C. nodosa}.
Chlorophyta associated proteins comprised 5.2 ± 0.06 \si{\percent} of
all identified proteins in \emph{C. cylindracea}, contributing 19.2 ±
1.5 \si{\percent} to the total protein abundance.

\hypertarget{discussion}{%
\subsection{Discussion}\label{discussion}}

To test whether the developed DNA and protein isolation protocols
efficiently detach microbes from the macrophyte surface we selected
\emph{C. nodosa} and \emph{C. cylindracea} as representatives of
seagrass and macroalgal species. These species differ morphologically.
While \emph{C. nodosa} leaves are flat, \emph{C. cylindracea} thalli are
characterized by an uneven surface (Kuo and den Hartog, 2001; Verlaque
\emph{et al.}, 2003). The developed protocol led to an almost complete
removal of epiphytic cells from the surfaces of both species comparable
to the result of Burke \emph{et al.} (2009), indicating that structural
differences do not impact the removal efficiency. In addition, isolation
protocols were tested in two contrasting seasons, as it is known that
macrophytes are harboring more algal epiphytes during autumn and winter
(Reyes and Sansón, 2001). No differences in the removal efficiency was
observed between seasons suggesting that these protocols can be used on
macrophyte samples retrieved throughout the year. Also, no removal
differences were observed on samples derived from the same host but from
different locations.

Successful amplification and sequencing of the V4 region of the 16S rRNA
gene proved that the isolated DNA can be used to estimate the microbial
epiphytic diversity. Taxonomic groups detected in this step can also be
often found in epiphytic communities associated with other macrophytes
(Burke and Thomas \emph{et al.}, 2011; Morrissey \emph{et al.}, 2019). A
problem often encountered in studies focusing on epiphytic communities
is the presence of large proportions of chloroplast 16S rRNA sequences
in the pool of amplified molecules, especially if the epiphytic DNA was
isolated without prior selection (Staufenberger \emph{et al.}, 2008).
These sequences can derive from host chloroplasts or from eukaryotic
epiphytic chloroplast DNA. Although the proportion of obtained
chloroplast 16S rRNA sequences in our samples was substantial, they
derived almost exclusively from eukaryotic epiphytes. High proportion of
chloroplast 16S rRNA sequences in studies applying selective procedures
that include direct cellular lysis on host surfaces were observed before
(Michelou \emph{et al.}, 2013). It is possible that chloroplast-specific
sequences even in these studies originated from eukaryotic epiphytic
cells and not from host chloroplasts. Indeed, it is common during 16S
rRNA profiling of pelagic microbial communities to observe high
proportions of chloroplast sequences (Gilbert \emph{et al.}, 2009;
Korlević \emph{et al.}, 2016). In addition, a very low proportion of
chloroplast 16S rRNA sequences in samples originating from \emph{C.
cylindracea} in comparison to \emph{C. nodosa} could be explained by the
presence of three introns in the gene for 16S rRNA in some members of
the genus \emph{Caulerpa} that could hamper the amplification process
(Lam and Lopez-Bautista, 2016).

High quality DNA is also needed for metagenomics. The obtained number of
metagenomic sequences and assembly statistics were comparable to
metagenomes and metatranscriptomes derived from similar surface
associated communities (Crump \emph{et al.}, 2018; Cúcio \emph{et al.},
2018). In addition, functional annotation of predicted coding sequences
to COG functional categories showed that the obtained metagenomes can be
used to determine the metabolic capacity of surface associated
communities (Leary \emph{et al.}, 2014; Cúcio \emph{et al.}, 2018). The
proportion of coding sequences, including their RPKM, originating from
\emph{C. nodosa} metagenomes and classified as Streptophyta was low
indicating that the isolation procedure was specific for epiphytic
cells. DNA samples isolated from the surface of \emph{C. cylindracea}
exhibited a low proportion of Chlorophyta coding sequences, however,
their RPKM was higher than in the samples originating from \emph{C.
nodosa}. One of the reasons for this elevated RPKM of Chlorophyta
sequences in \emph{C. cylindracea} could be the differences in the
tissue structure between these two host species. While \emph{C. nodosa}
leaves are composed of individual cells, the thallus of \emph{C.
cylindracea} is, like in other siphonous algal species, composed of a
single large multinucleate cell (Coneva and Chitwood, 2015). The absence
of individual cells in \emph{C. cylindracea} could cause a leakage of
genetic material into the extraction buffer causing an elevated presence
of host sequences in the samples for metagenome analyses.

To obtain insight into the metabolic status of uncultivated prokaryotes,
a metaproteomic approach is required (Saito \emph{et al.}, 2019). The
applied protocol for epiphytic protein isolation followed by a
metaproteomic analysis identified between 14,219 and 16,449 proteins,
which is higher than previously reported for e.g.~soils (Chourey
\emph{et al.}, 2010; Hultman \emph{et al.}, 2015), seawater (Williams
\emph{et al.}, 2012) and biofilms (Leary \emph{et al.}, 2014). The
functional annotation of identified proteins into COG functional
categories showed that the protein isolation protocol can be used to
assess the metabolic status of the epiphytic community (Leary \emph{et
al.}, 2014). Similar to the results of the metagenomic analysis, the
number and abundance of identified proteins affiliated to Streptophyta
in \emph{C. nodosa} samples were low, indicating that the procedure is
selective for epiphytic cell proteins. In addition, a higher number and
abundance of identified proteins associated with Chlorophyta were
observed in \emph{C. cylindracea} samples. The cause of this elevated
presence of Chlorophyta-associated proteins can be, similar to the DNA
isolation protocol, explained by the absence of individual cells in this
siphonous alga (Coneva and Chitwood, 2015).

In conclusion, the developed protocols for DNA and protein isolation
from macrophyte surfaces almost completely remove the epiphytic
community from both, \emph{C. nodosa} and \emph{C. cylindracea}, in
different seasons. Also, the obtained DNA and proteins are suitable for
16S rRNA sequencing, metagenomics and metaproteomics analyses while the
obtained material contains low quantities of host DNA and proteins
making the protocols specific for epiphytes. Furthermore, the protocols
are based on universally available laboratory chemicals hence, making
them widely applicable.

\newpage

\hypertarget{experimental-procedures}{%
\subsection{Experimental procedures}\label{experimental-procedures}}

\hypertarget{sampling}{%
\subsubsection{Sampling}\label{sampling}}

Leaves of \emph{C. nodosa} were sampled in a \emph{C. nodosa} meadow in
the Bay of Saline, northern Adriatic Sea (\ang{45;7;5} N, \ang{13;37;20}
E) and in a \emph{C. nodosa} meadow invaded by \emph{C. cylindracea} in
the Bay of Funtana, northern Adriatic Sea (\ang{45;10;39} N,
\ang{13;35;42} E). Thalli of \emph{C. cylindracea} were sampled in the
same \emph{C. nodosa} invaded meadow in the Bay of Funtana and at a
locality of only \emph{C. cylindracea} located in the proximity of the
invaded meadow. Leaves and thalli for 16S rRNA analysis, metagenomics
and metaproteomics were collected in two contrasting seasons, on 4
December 2017 (16S rRNA analysis and metaproteomics), 14 December 2017
(metagenomics) and 18 June 2018 (16S rRNA analysis, metagenomics and
metaproteomics). During spring 2018, the \emph{C. nodosa} meadow in the
Bay of Saline decayed to an extent that no leaves could be retrieved
(Najdek \emph{et al.}, 2020). In addition, as not enough DNA for both
metagenomic and 16S RNA analysis were obtained during the sampling on 4
December 2017, an additional sampling on 14 December 2017 was carried
out in the Bay of Funtana. Leaves and thalli were collected by diving
and transported to the laboratory in containers placed on ice and filled
with seawater from this site. Upon arrival to the laboratory, \emph{C.
nodosa} leaves were cut into sections of 1 -- 2 \si{\cm}, while \emph{C.
cylindracea} thalli were cut into 5 -- 8 \si{\cm} long sections. Leaves
and thalli were washed three times with sterile artificial seawater
(ASW) to remove loosely attached microbial cells.

\hypertarget{dna-isolation}{%
\subsubsection{DNA isolation}\label{dna-isolation}}

The DNA was isolated according to the protocol for isolation from
filters described in Massana \emph{et al.} (1997). This protocol was
modified and adapted for microbial DNA isolation from macrophyte
surfaces as described below. Five (5) \si{\ml} of lysis buffer (40
\si{\milli\Molar} EDTA, 50 \si{\milli\Molar} Tris-HCl, 0.75 \si{\Molar}
sucrose; pH 8.3) was added to 1 \si{\g} wet weight of leaves or 2
\si{\g} wet-weight of thalli. For every sample, duplicate isolations
were performed. Lysozyme was added (final concentration 1
\si{\mg\per\ml}) and the mixture was incubated at 37 \si{\degreeCelsius}
for 30 \si{\minute}. Subsequently, proteinase K (final concentration 0.5
\si{\mg\per\ml}) and SDS (final concentration 1 \si{\percent}) were
added and the samples were incubated at 55 \si{\degreeCelsius} for 2
\si{\hour}. Following the incubation, tubes were vortexed for 10
\si{\minute} and the mixture containing lysed epiphytic cells was
separated from host leaves or thalli by transferring the solution into a
clean tube. The lysate was extracted twice with a mixture of
phenol:chloroform:isoamyl alcohol (25:24:1; pH 8) and once with
chloroform:isoamyl alcohol (24:1). After each addition of an organic
solvent mixture, tubes were slightly vortexed and centrifuged at 4,500 ×
g for 10 \si{\minute}. Following each centrifugation the aqueous phases
were retrieved. After the final extraction 1/10 of chilled 3 \si{\Molar}
sodium acetate (pH 5.2) was added. DNA was precipitated by adding 1
volume of chilled isopropanol, incubating the mixtures overnight at
\num{-20} \si{\degreeCelsius} and centrifuging at 16,000 × g and 4
\si{\degreeCelsius} for 20 \si{\minute}. The pellet was washed twice
with 1 \si{\ml} of chilled 70 \si{\percent} ethanol and centrifuged
after each washing step at 20,000 × g and 4 \si{\degreeCelsius} for 10
\si{\minute}. After the first washing step duplicate pellets from the
same sample were pooled and transferred to a clean 1.5 \si{\ml} tube.
The dried pellet was resuspended in 100 \si{\ul} of deionized water.

\hypertarget{illumina-16s-rrna-sequencing}{%
\subsubsection{Illumina 16S rRNA
sequencing}\label{illumina-16s-rrna-sequencing}}

An aliquot of isolated DNA was treated with RNase A (final concentration
200 \si{\ug\per\ml}) for 2 \si{\hour} at 37 \si{\degreeCelsius}. The DNA
concentration was determined using the Quant-iT PicoGreen dsDNA Assay
Kit (Invitrogen, USA) according to the manufacturer's instructions and
diluted to 1 \si{\ng\per\ul}. The V4 region of the 16S rRNA gene was
amplified using a two-step PCR procedure. In the first PCR the 515F
(5'-GTGYCAGCMGCCGCGGTAA-3') and 806R (5'-GGACTACNVGGGTWTCTAAT-3')
primers from the Earth Microbiome Project
(\url{http://press.igsb.anl.gov/earthmicrobiome/protocols-and-standards/16s/})
were used to amplify the target region (Caporaso \emph{et al.}, 2012;
Apprill \emph{et al.}, 2015; Parada \emph{et al.}, 2016). These primers
contained on their 5' end a tagged sequence. Each sample was amplified
in four parallel 25 \si{\ul} reactions of which each contained 1 × Q5
Reaction Buffer, 0.2 \si{\milli\Molar} of dNTPmix, 0.7 \si{\mg\per\ml}
BSA (Bovine Serum Albumin), 0.2 \si{\micro\Molar} of forward and reverse
primers, 0.5 U of Q5 High-Fidelity DNA Polymerase (New England Biolabs,
USA) and 5 \si{\ng} of DNA template. Cycling conditions were: initial
denaturation at 94 \si{\degreeCelsius} for 3 \si{\minute}, 20 cycles of
denaturation at 94 \si{\degreeCelsius} for 45 \si{\s}, annealing at 50
\si{\degreeCelsius} for 60 \si{\s} and elongation at 72
\si{\degreeCelsius} for 90 \si{\s}, finalized by an elongation step at
72 \si{\degreeCelsius} for 10 \si{\minute}. The four parallel reactions
volumes were pooled and PCR products were purified using the GeneJET PCR
Purification Kit (Thermo Fisher Scientific, USA) according to the
manufacturer's instructions and following the protocol that included
isopropanol addition for better small DNA fragment yield. The column was
eluted in 30 \si{\ul} of deionized water. Purified PCR products were
sent for Illumina MiSeq sequencing (2 × 250 bp) at IMGM Laboratories,
Martinsried, Germany. Before sequencing at IMGM, the second PCR
amplification of the two-step PCR procedure was performed using primers
targeting the tagged region incorporated in the first PCR. In addition,
these primers contained adapter and sample-specific index sequences. The
second PCR was carried out for 8 cycles. Beside samples, a positive and
negative control were sequenced. A negative control was comprised of
four parallel PCR reactions without DNA template, while for a positive
control a mock community composed of evenly mixed DNA material
originating from 20 bacterial strains (ATCC MSA-1002, ATCC, USA) was
used. Partial 16S rRNA sequences obtained in this study have been
deposited in the European Nucleotide Archive (ENA) at EMBL-EBI under
accession numbers SAMEA6786270, SAMEA6648792 -- SAMEA6648794,
SAMEA6648809 -- SAMEA6648811.

Obtained sequences were analyzed on the computer cluster Isabella
(University Computing Center, University of Zagreb) using mothur
(version 1.43.0) (Schloss \emph{et al.}, 2009) according to the MiSeq
Standard Operating Procedure (MiSeq SOP;
\url{https://mothur.org/wiki/MiSeq_SOP}) (Kozich \emph{et al.}, 2013)
and recommendations given from the Riffomonas project to enhance data
reproducibility (\url{http://www.riffomonas.org/}). For alignment and
classification of sequences the SILVA SSU Ref NR 99 database (release
138; \url{http://www.arb-silva.de}) was used (Quast \emph{et al.}, 2013;
Yilmaz \emph{et al.}, 2014). Sequences classified as chloroplasts by
SILVA were exported and reclassified using mothur and the RDP
(\url{http://rdp.cme.msu.edu/}) training set (version 16) reference
files adapted for mothur (Cole \emph{et al.}, 2014). In comparison to
SILVA, RDP allows a more detailed classification of chloroplast
sequences. Based on the ATCC MSA-1002 mock community included in the
analysis a sequencing error rate of 0.009 \si{\percent} was determined,
which is in line with previously reported values for next-generation
sequencing data (Kozich \emph{et al.}, 2013; Schloss \emph{et al.},
2016). In addition, the negative control processed together with the
samples yielded only 2 sequences after sequence quality curation.

\hypertarget{metagenomics}{%
\subsubsection{Metagenomics}\label{metagenomics}}

Four DNA samples were selected and sent on dry ice to IMGM Laboratories,
Martinsried, Germany for metagenomic sequencing. DNA was purified using
AMPure XP Beads (Beckman Coulter, USA) applying a bead:DNA ratio of 1:1
(v/v), quantified with a Qubit dsDNA HS Assay Kit (Thermo Fisher
Scientific, USA) and check for integrity on a 1 \si{\percent} agarose
gel. Metagenomic sequencing libraries were generated from 300 \si{\ng}
of input DNA using a NEBNext Ultra II FS DNA Library Prep Kit for
Illumina (New England Biolabs, USA) according to the manufacturer's
instructions. Fragments were selected (500 -- 700 bp) using AMPure XP
Beads, PCR enriched for 3 -- 5 cycles and quality controlled. Libraries
generated from different DNA samples were pooled and sequenced on an
Illumina NovaSeq 6000 sequencing system (2 × 250 bp).

Obtained sequences were analyzed on the Life Science Compute Cluster
(LiSC) (CUBE -- Computational Systems Biology, University of Vienna).
Individual sequences were assembled using MEGAHIT (version 1.1.2) (Li
\emph{et al.}, 2015) under default settings. Putative genes were
predicted from contings longer than 200 bp using Prodigal (version
2.6.3) (Hyatt \emph{et al.}, 2010) in metagenome mode (-p meta).
Abundances of predicted genes were expressed as Reads Per Kilobase
Million (RPKM) and calculated using the BWA algorithm (version 0.7.16a)
(Li and Durbin, 2009). All predicted genes were functionally annotated
using the eggNOG-mapper (Huerta-Cepas \emph{et al.}, 2017) and eggNOG
database (version 5.0) (Huerta-Cepas \emph{et al.}, 2019). Sequence
taxonomy classification was determined using the lowest common ancestor
algorithm adapted from DIAMOND (version 0.8.36) (Buchfink \emph{et al.},
2015) and by searching against the NCBI non-redundant database (NR). To
determine the phylogeny, the top 10 \si{\percent} hits with an e-value
\textless{} 1 × 10\textsuperscript{-5} were used (-\/-top 10). Sequence
renaming, coverage information computing and metagenomic statistics
calculations were performed using software tools from BBTools
(\url{https://jgi.doe.gov/data-and-tools/bbtools}). Metagenomic
sequences obtained in this study have been deposited in the European
Nucleotide Archive (ENA) at EMBL-EBI under accession numbers
SAMEA6648795, SAMEA6648797, SAMEA6648809 and SAMEA6648811.

\hypertarget{protein-isolation}{%
\subsubsection{Protein isolation}\label{protein-isolation}}

Proteins were isolated according to the protocol for protein isolation
from soil described in Chourey \emph{et al.} (2010) and modified by
Hultman \emph{et al.} (2015). This protocol was further modified and
adapted for microbial protein isolation from macrophyte surfaces as
described below. Twenty (20) \si{\ml} of protein extraction buffer (4
\si{\percent} SDS, 100 \si{\milli\Molar} Tris-HCl {[}pH 8.0{]}) was
added to 5 \si{\g} wet weight of leaves or 10 \si{\g} wet weight of
thalli. The mixture was incubated in boiling water for 5 \si{\minute},
vortexed for 10 \si{\minute} and incubated again in boiling water for 5
\si{\minute}. After a brief vortex, the lysate was transferred to a
clean tube separating the host leaves or thalli from the mixture
containing lysed epiphytic cells. Dithiothreitol (DTT; final
concentration 24 \si{\milli\Molar}) was added and proteins were
precipitated with chilled 100 \si{\percent} trichloroacetic acid (TCA;
final concentration 20 \si{\percent}) overnight at \num{-20} °C.
Precipitated proteins were centrifuged at 10,000 × g and 4
\si{\degreeCelsius} for 40 \si{\minute}. The obtained protein pellet was
washed three times with chilled acetone. During the first washing step
the pellet was transferred to a clean 1.5 \si{\ml} tube. After each
washing step, samples were centrifuged at 20,000 × g and 4
\si{\degreeCelsius} for 5 \si{\minute}. Dried pellets were stored at
\num{-80} °C until further analysis.

\hypertarget{metaproteomics}{%
\subsubsection{Metaproteomics}\label{metaproteomics}}

Isolated proteins were whole trypsin digested using the FASP
(filter-aided sample preparation) Protein Digestion Kit (Expedeon, UK)
according to the manufacturer's instructions with small modifications
(Wiśniewski \emph{et al.}, 2009). Prior to loading the solution onto the
column, protein pellets were solubilized in a urea sample buffer
included in the kit amended with DTT (final concentration 100
\si{\milli\Molar}) for 45 \si{\minute} at room temperature and
centrifuged at 20,000 × g for 2 -- 5 \si{\minute} at room temperature to
remove larger particles. The first washing step after protein solution
loading was repeated twice. In addition, the centrifugation steps were
prolonged if the column was clogged. Trypsin digestion was performed on
column filters at 37 \si{\degreeCelsius} overnight for 18 \si{\hour}.
The final filtrate containing peptides was acidified with 1
\si{\percent} (final concentration) trifluoroacetic acid (TFA), freezed
at \num{-80} \si{\degreeCelsius}, lyophilized and sent to VIME -- Vienna
Metabolomics Center (University of Vienna) for metaproteomic analysis.
Peptides were resuspended in 1 \si{\percent} (final concentration) TFA,
desalted using the Pierce C18 Tips (Thermo Fisher Scientific, USA)
according to the manufacturer's instructions and sequenced on a Q
Exactive Hybrid Quadrupole-Orbitrap Mass Spectrometer (Thermo Fisher
Scientific, USA). Obtained MS/MS spectra were searched against a protein
database composed of combined sequenced metagenomes using SEQUEST-HT
engines and validated with Percolator in Proteome Discoverer 2.1 (Thermo
Fisher Scientific, USA). The target-decoy approach was used to reduce
the probability of false peptide identification. Results whose false
discovery rate at the peptide level was \SI{< 1}{\percent} were kept.
For protein identification a minimum of two peptides and one unique
peptide were required. For protein quantification, a chromatographic
peak area-based free quantitative method was applied.

\hypertarget{data-processing-and-visualization}{%
\subsubsection{Data processing and
visualization}\label{data-processing-and-visualization}}

Processing and visualization of 16S rRNA, metagenomic and metaproteomic
data were done using R (version 3.6.0) (R Core Team, 2019), package
tidyverse (version 1.3.0) (Wickham \emph{et al.}, 2019) and multiple
other packages (Xie, 2014, 2015, 2019, 2020; Neuwirth, 2014; Xie
\emph{et al.}, 2018; Wilke, 2018; Allaire \emph{et al.}, 2019; Zhu,
2019; Bengtsson, 2020). The detailed analysis procedure including the R
Markdown file for this paper are available as a GitHub repository
(\url{https://github.com/MicrobesRovinj/Korlevic_SelectiveRemoval_EnvironMicrobiol_2020}).

\hypertarget{confocal-microscopy}{%
\subsubsection{Confocal microscopy}\label{confocal-microscopy}}

Host leaves and thalli from DNA and protein isolation steps were washed
seven times in deionized water and fixed with formaldehyde (final
concentration \textasciitilde{} 3 \si{\percent}). In addition,
nontreated leaves and thalli, washed three times in ASW to remove
loosely attached microbial cells, were fixed in the same concentration
of formaldehyde and used as a positive control. For-long term storage,
fixed leaves and thalli were immersed in a mixture of phosphate-buffered
saline (PBS) and ethanol (1:1) and stored at \num{-20}
\si{\degreeCelsius}. Treated and untreated segments of leaves and thalli
were stained in a 2 × solution of SYBR Green I and examined under a
Leica TCS SP8 X FLIM confocal microscope (Leica Microsystems, Germany).

\hypertarget{acknowledgements}{%
\subsection{Acknowledgements}\label{acknowledgements}}

This work was funded by the Croatian Science Foundation through the
MICRO-SEAGRASS project (project number IP-2016-06-7118). ZZ and GJH were
supported by the Austrian Science Fund (FWF) project ARTEMIS (project
number P28781-B21). We would like to thank Margareta Buterer for
technical support, Paolo Paliaga for help during sampling, Lucija Horvat
for technical support in confocal microscopy, Dušica Vujaklija for help
in accessing the confocal microscope and Sonja Tischler for technical
support in mass spectrometry. In addition, we would like to thank the
University Computing Center of the University of Zagreb for access to
the computer cluster Isabella and CUBE -- Computational Systems Biology
of the University of Vienna for access to the Life Science Compute
Cluster (LiSC).

\newpage

\hypertarget{references}{%
\subsection{References}\label{references}}

\hypertarget{refs}{}
\leavevmode\hypertarget{ref-Allaire2019}{}%
Allaire, J.J., Xie, Y., McPherson, J., Luraschi, J., Ushey, K., Atkins,
A., et al. (2019) rmarkdown: Dynamic documents for R.

\leavevmode\hypertarget{ref-Amann1995}{}%
Amann, R.I., Ludwig, W., and Schleifer, K.H. (1995) Phylogenetic
identification and \emph{in situ} detection of individual microbial
cells without cultivation. \emph{Microbiol Rev} \textbf{59}: 143--169.

\leavevmode\hypertarget{ref-Andersson2010}{}%
Andersson, A.F., Riemann, L., and Bertilsson, S. (2010) Pyrosequencing
reveals contrasting seasonal dynamics of taxa within Baltic Sea
bacterioplankton communities. \emph{ISME J} \textbf{4}: 171--181.

\leavevmode\hypertarget{ref-Apprill2015}{}%
Apprill, A., McNally, S., Parsons, R., and Weber, L. (2015) Minor
revision to V4 region SSU rRNA 806R gene primer greatly increases
detection of SAR11 bacterioplankton. \emph{Aquat Microb Ecol}
\textbf{75}: 129--137.

\leavevmode\hypertarget{ref-Bengtsson2020}{}%
Bengtsson, H. (2020) matrixStats: Functions that apply to rows and
columns of matrices (and to vectors).

\leavevmode\hypertarget{ref-Buchfink2015}{}%
Buchfink, B., Xie, C., and Huson, D.H. (2015) Fast and sensitive protein
alignment using DIAMOND. \emph{Nat Methods} \textbf{12}: 59--60.

\leavevmode\hypertarget{ref-Burke2009}{}%
Burke, C., Kjelleberg, S., and Thomas, T. (2009) Selective extraction of
bacterial DNA from the surfaces of macroalgae. \emph{Appl Environ
Microbiol} \textbf{75}: 252--256.

\leavevmode\hypertarget{ref-Burke2011}{}%
Burke, C., Steinberg, P., Rusch, D., Kjelleberg, S., and Thomas, T.
(2011) Bacterial community assembly based on functional genes rather
than species. \emph{Proc Natl Acad Sci U S A} \textbf{108}:
14288--14293.

\leavevmode\hypertarget{ref-Burke2011a}{}%
Burke, C., Thomas, T., Lewis, M., Steinberg, P., and Kjelleberg, S.
(2011) Composition, uniqueness and variability of the epiphytic
bacterial community of the green alga \emph{Ulva australis}. \emph{ISME
J} \textbf{5}: 590--600.

\leavevmode\hypertarget{ref-Cai2014}{}%
Cai, X., Gao, G., Yang, J., Tang, X., Dai, J., Chen, D., and Song, Y.
(2014) An ultrasonic method for separation of epiphytic microbes from
freshwater submerged macrophytes. \emph{J Basic Microbiol} \textbf{54}:
758--761.

\leavevmode\hypertarget{ref-Caporaso2012}{}%
Caporaso, J.G., Lauber, C.L., Walters, W.A., Berg-Lyons, D., Huntley,
J., Fierer, N., et al. (2012) Ultra-high-throughput microbial community
analysis on the Illumina HiSeq and MiSeq platforms. \emph{ISME J}
\textbf{6}: 1621--1624.

\leavevmode\hypertarget{ref-Chourey2010}{}%
Chourey, K., Jansson, J., VerBerkmoes, N., Shah, M., Chavarria, K.L.,
Tom, L.M., et al. (2010) Direct cellular lysis/protein extraction
protocol for soil metaproteomics. \emph{J Proteome Res} \textbf{9}:
6615--6622.

\leavevmode\hypertarget{ref-Cole2014}{}%
Cole, J.R., Wang, Q., Fish, J.A., Chai, B., McGarrell, D.M., Sun, Y., et
al. (2014) Ribosomal Database Project: Data and tools for high
throughput rRNA analysis. \emph{Nucleic Acids Res} \textbf{42}:
D633--D642.

\leavevmode\hypertarget{ref-Coneva2015}{}%
Coneva, V. and Chitwood, D.H. (2015) Plant architecture without
multicellularity: Quandaries over patterning and the soma-germline
divide in siphonous algae. \emph{Front Plant Sci} \textbf{6}: 287.

\leavevmode\hypertarget{ref-Crump2018}{}%
Crump, B.C., Wojahn, J.M., Tomas, F., and Mueller, R.S. (2018)
Metatranscriptomics and amplicon sequencing reveal mutualisms in
seagrass microbiomes. \emph{Front Microbiol} \textbf{9}: 388.

\leavevmode\hypertarget{ref-Cucio2018}{}%
Cúcio, C., Overmars, L., Engelen, A.H., and Muyzer, G. (2018)
Metagenomic analysis shows the presence of bacteria related to
free-living forms of sulfur-oxidizing chemolithoautotrophic symbionts in
the rhizosphere of the seagrass \emph{Zostera marina}. \emph{Front Mar
Sci} \textbf{5}: 171.

\leavevmode\hypertarget{ref-Egan2013}{}%
Egan, S., Harder, T., Burke, C., Steinberg, P., Kjelleberg, S., and
Thomas, T. (2013) The seaweed holobiont: Understanding seaweed-bacteria
interactions. \emph{FEMS Microbiol Rev} \textbf{37}: 462--476.

\leavevmode\hypertarget{ref-Gilbert2009}{}%
Gilbert, J.A., Field, D., Swift, P., Newbold, L., Oliver, A., Smyth, T.,
et al. (2009) The seasonal structure of microbial communities in the
Western English Channel. \emph{Environ Microbiol} \textbf{11}:
3132--3139.

\leavevmode\hypertarget{ref-Gross2003}{}%
Gross, E.M., Feldbaum, C., and Graf, A. (2003) Epiphyte biomass and
elemental composition on submersed macrophytes in shallow eutrophic
lakes. \emph{Hydrobiologia} \textbf{506-509}: 559--565.

\leavevmode\hypertarget{ref-Huerta-Cepas2017}{}%
Huerta-Cepas, J., Forslund, K., Coelho, L.P., Szklarczyk, D., Jensen,
L.J., von Mering, C., and Bork, P. (2017) Fast genome-wide functional
annotation through orthology assignment by eggNOG-mapper. \emph{Mol Biol
Evol} \textbf{34}: 2115--2122.

\leavevmode\hypertarget{ref-Huerta-Cepas2019}{}%
Huerta-Cepas, J., Szklarczyk, D., Heller, D., Hernández-Plaza, A.,
Forslund, S.K., Cook, H., et al. (2019) eggNOG 5.0: A hierarchical,
functionally and phylogenetically annotated orthology resource based on
5090 organisms and 2502 viruses. \emph{Nucleic Acids Res} \textbf{47}:
D309--D314.

\leavevmode\hypertarget{ref-Hultman2015}{}%
Hultman, J., Waldrop, M.P., Mackelprang, R., David, M.M., McFarland, J.,
Blazewicz, S.J., et al. (2015) Multi-omics of permafrost, active layer
and thermokarst bog soil microbiomes. \emph{Nature} \textbf{521}:
208--212.

\leavevmode\hypertarget{ref-Hyatt2010}{}%
Hyatt, D., Chen, G.-L., LoCascio, P.F., Land, M.L., Larimer, F.W., and
Hauser, L.J. (2010) Prodigal: Prokaryotic gene recognition and
translation initiation site identification. \emph{BMC Bioinformatics}
\textbf{11}: 119.

\leavevmode\hypertarget{ref-Jiang2015}{}%
Jiang, Y.-F., Ling, J., Dong, J.-D., Chen, B., Zhang, Y.-Y., Zhang,
Y.-Z., and Wang, Y.-S. (2015) Illumina-based analysis the microbial
diversity associated with \emph{Thalassia hemprichii} in Xincun Bay,
South China Sea. \emph{Ecotoxicology} \textbf{24}: 1548--56.

\leavevmode\hypertarget{ref-Korlevic2016}{}%
Korlević, M., Šupraha, L., Ljubešić, Z., Henderiks, J., Ciglenečki, I.,
Dautović, J., and Orlić, S. (2016) Bacterial diversity across a highly
stratified ecosystem: A salt-wedge Mediterranean estuary. \emph{Syst
Appl Microbiol} \textbf{39}: 398--408.

\leavevmode\hypertarget{ref-Kozich2013}{}%
Kozich, J.J., Westcott, S.L., Baxter, N.T., Highlander, S.K., and
Schloss, P.D. (2013) Development of a dual-index sequencing strategy and
curation pipeline for analyzing amplicon sequence data on the MiSeq
Illumina sequencing platform. \emph{Appl Environ Microbiol} \textbf{79}:
5112--5120.

\leavevmode\hypertarget{ref-Kuo2001}{}%
Kuo, J. and den Hartog, C. (2001) Seagrass taxonomy and identification
key. In \emph{Global Seagrass Research Methods}. Short, F.T. and Coles,
R.G. (eds). Amsterdam: Elsevier Science B.V., pp. 31--58.

\leavevmode\hypertarget{ref-Lam2016}{}%
Lam, D.W. and Lopez-Bautista, J.M. (2016) Complete chloroplast genome
for \emph{Caulerpa racemosa} (Bryopsidales, Chlorophyta) and comparative
analyses of siphonous green seaweed plastomes. \emph{Cymbella}
\textbf{2}: 23--32.

\leavevmode\hypertarget{ref-Leary2014}{}%
Leary, D.H., Li, R.W., Hamdan, L.J., IV, W.J.H., Lebedev, N., Wang, Z.,
et al. (2014) Integrated metagenomic and metaproteomic analyses of
marine biofilm communities. \emph{Biofouling} \textbf{30}: 1211--1223.

\leavevmode\hypertarget{ref-Li2015}{}%
Li, D., Liu, C.-M., Luo, R., Sadakane, K., and Lam, T.-W. (2015)
MEGAHIT: An ultra-fast single-node solution for large and complex
metagenomics assembly via succinct de Bruijn graph.
\emph{Bioinformatics} \textbf{31}: 1674--1676.

\leavevmode\hypertarget{ref-Li2009a}{}%
Li, H. and Durbin, R. (2009) Fast and accurate short read alignment with
BurrowsWheeler transform. \emph{Bioinformatics} \textbf{25}: 1754--1760.

\leavevmode\hypertarget{ref-Longford2007}{}%
Longford, S., Tujula, N., Crocetti, G., Holmes, A., Holmström, C.,
Kjelleberg, S., et al. (2007) Comparisons of diversity of bacterial
communities associated with three sessile marine eukaryotes. \emph{Aquat
Microb Ecol} \textbf{48}: 217--229.

\leavevmode\hypertarget{ref-Massana1997}{}%
Massana, R., Murray, A.E., Preston, C.M., and DeLong, E.F. (1997)
Vertical distribution and phylogenetic characterization of marine
planktonic \emph{Archaea} in the Santa Barbara Channel. \emph{Appl
Environ Microbiol} \textbf{63}: 50--56.

\leavevmode\hypertarget{ref-Michelou2013}{}%
Michelou, V.K., Caporaso, J.G., Knight, R., and Palumbi, S.R. (2013) The
ecology of microbial communities associated with \emph{Macrocystis
pyrifera}. \emph{PloS One} \textbf{8}: e67480.

\leavevmode\hypertarget{ref-Morrissey2019}{}%
Morrissey, K.L., Çavas, L., Willems, A., and De Clerck, O. (2019)
Disentangling the influence of environment, host specificity and thallus
differentiation on bacterial communities in siphonous green seaweeds.
\emph{Front Microbiol} \textbf{10}: 717.

\leavevmode\hypertarget{ref-Najdek2020}{}%
Najdek, M., Korlević, M., Paliaga, P., Markovski, M., Ivančić, I.,
Iveša, L., et al. (2020) Dynamics of environmental conditions during the
decline of a \emph{Cymodocea nodosa} meadow. \emph{Biogeosciences}
\textbf{17}: 3299--3315.

\leavevmode\hypertarget{ref-Neuwirth2014}{}%
Neuwirth, E. (2014) RColorBrewer: ColorBrewer palettes.

\leavevmode\hypertarget{ref-Noges2010}{}%
Nõges, T., Luup, H., and Feldmann, T. (2010) Primary production of
aquatic macrophytes and their epiphytes in two shallow lakes (Peipsi and
Võrtsjärv) in Estonia. \emph{Aquat Ecol} \textbf{44}: 83--92.

\leavevmode\hypertarget{ref-Parada2016}{}%
Parada, A.E., Needham, D.M., and Fuhrman, J.A. (2016) Every base
matters: Assessing small subunit rRNA primers for marine microbiomes
with mock communities, time series and global field samples.
\emph{Environ Microbiol} \textbf{18}: 1403--1414.

\leavevmode\hypertarget{ref-Quast2013}{}%
Quast, C., Pruesse, E., Yilmaz, P., Gerken, J., Schweer, T., Yarza, P.,
et al. (2013) The SILVA ribosomal RNA gene database project: Improved
data processing and web-based tools. \emph{Nucleic Acids Res}
\textbf{41}: D590--D596.

\leavevmode\hypertarget{ref-RCoreTeam2019}{}%
R Core Team (2019) R: A language and environment for statistical
computing, Vienna, Austria: R Foundation for Statistical Computing.

\leavevmode\hypertarget{ref-Reyes2001}{}%
Reyes, J. and Sansón, M. (2001) Biomass and production of the epiphytes
on the leaves of \emph{Cymodocea nodosa} in the Canary Islands.
\emph{Bot Mar} \textbf{44}: 307--313.

\leavevmode\hypertarget{ref-Richter-Heitmann2016}{}%
Richter-Heitmann, T., Eickhorst, T., Knauth, S., Friedrich, M.W., and
Schmidt, H. (2016) Evaluation of strategies to separate root-associated
microbial communities: A crucial choice in rhizobiome research.
\emph{Front Microbiol} \textbf{7}: 773.

\leavevmode\hypertarget{ref-Saito2019}{}%
Saito, M.A., Bertrand, E.M., Duffy, M.E., Gaylord, D.A., Held, N.A.,
Hervey, W.J., et al. (2019) Progress and challenges in ocean
metaproteomics and proposed best practices for data sharing. \emph{J
Proteome Res} \textbf{18}: 1461--1476.

\leavevmode\hypertarget{ref-Schloss2016}{}%
Schloss, P.D., Jenior, M.L., Koumpouras, C.C., Westcott, S.L., and
Highlander, S.K. (2016) Sequencing 16S rRNA gene fragments using the
PacBio SMRT DNA sequencing system. \emph{PeerJ} \textbf{4}: e1869.

\leavevmode\hypertarget{ref-Schloss2009}{}%
Schloss, P.D., Westcott, S.L., Ryabin, T., Hall, J.R., Hartmann, M.,
Hollister, E.B., et al. (2009) Introducing mothur: Open-source,
platform-independent, community-supported software for describing and
comparing microbial communities. \emph{Appl Environ Microbiol}
\textbf{75}: 7537--7541.

\leavevmode\hypertarget{ref-Staufenberger2008}{}%
Staufenberger, T., Thiel, V., Wiese, J., and Imhoff, J.F. (2008)
Phylogenetic analysis of bacteria associated with \emph{Laminaria
saccharina}. \emph{FEMS Microbiol Ecol} \textbf{64}: 65--77.

\leavevmode\hypertarget{ref-Su2012}{}%
Su, C., Lei, L., Duan, Y., Zhang, K.-Q., and Yang, J. (2012)
Culture-independent methods for studying environmental microorganisms:
Methods, application, and perspective. \emph{Appl Microbiol Biotechnol}
\textbf{93}: 993--1003.

\leavevmode\hypertarget{ref-Uku2007}{}%
Uku, J., Björk, M., Bergman, B., and Díez, B. (2007) Characterization
and comparison of prokaryotic epiphytes associated with three East
African seagrasses. \emph{J Phycol} \textbf{43}: 768--779.

\leavevmode\hypertarget{ref-Verlaque2003}{}%
Verlaque, M., Durand, C., Huisman, J.M., Boudouresque, C.-F., and Le
Parco, Y. (2003) On the identity and origin of the Mediterranean
invasive \emph{Caulerpa} \emph{Racemosa} (Caulerpales, Chlorophyta).
\emph{Eur J Phycol} \textbf{38}: 325--339.

\leavevmode\hypertarget{ref-Weidner1996}{}%
Weidner, S., Arnold, W., and Puhler, A. (1996) Diversity of uncultured
microorganisms associated with the seagrass \emph{Halophila stipulacea}
estimated by restriction fragment length polymorphism analysis of
PCR-amplified 16S rRNA genes. \emph{Appl Environ Microbiol} \textbf{62}:
766--771.

\leavevmode\hypertarget{ref-Wickham2019}{}%
Wickham, H., Averick, M., Bryan, J., Chang, W., McGowan, L.D., François,
R., et al. (2019) Welcome to the tidyverse. \emph{J Open Source Softw}
\textbf{4}: 1686.

\leavevmode\hypertarget{ref-Wilke2018}{}%
Wilke, C.O. (2018) cowplot: Streamlined plot theme and plot annotations
for 'ggplot2'.

\leavevmode\hypertarget{ref-Williams2012}{}%
Williams, T.J., Long, E., Evans, F., DeMaere, M.Z., Lauro, F.M.,
Raftery, M.J., et al. (2012) A metaproteomic assessment of winter and
summer bacterioplankton from Antarctic Peninsula coastal surface waters.
\emph{ISME J} \textbf{6}: 1883--1900.

\leavevmode\hypertarget{ref-Wisniewski2009}{}%
Wiśniewski, J.R., Zougman, A., Nagaraj, N., and Mann, M. (2009)
Universal sample preparation method for proteome analysis. \emph{Nat
Methods} \textbf{6}: 359--362.

\leavevmode\hypertarget{ref-Xie2015}{}%
Xie, Y. (2015) Dynamic Documents with R and knitr, 2nd ed. Boca Raton,
Florida: Chapman and Hall/CRC.

\leavevmode\hypertarget{ref-Xie2014}{}%
Xie, Y. (2014) knitr: A comprehensive tool for reproducible research in
R. In \emph{Implementing Reproducible Computational Research}. Stodden,
V., Leisch, F., and Peng, R.D. (eds). New York: Chapman and Hall/CRC,
pp. 3--32.

\leavevmode\hypertarget{ref-Xie2019a}{}%
Xie, Y. (2019) TinyTeX: A lightweight, cross-platform, and
easy-to-maintain LaTeX distribution based on TeX Live. \emph{TUGboat}
\textbf{40}: 30--32.

\leavevmode\hypertarget{ref-Xie2020}{}%
Xie, Y. (2020) tinytex: Helper functions to install and maintain 'TeX
Live', and compile 'LaTeX' documents.

\leavevmode\hypertarget{ref-Xie2018}{}%
Xie, Y., Allaire, J.J., and Grolemund, G. (2018) R Markdown: The
Definitive Guide, 1st ed. Boca Raton, Florida: Chapman and Hall/CRC.

\leavevmode\hypertarget{ref-Yilmaz2014}{}%
Yilmaz, P., Parfrey, L.W., Yarza, P., Gerken, J., Pruesse, E., Quast,
C., et al. (2014) The SILVA and "All-Species Living Tree Project (LTP)"
taxonomic frameworks. \emph{Nucleic Acids Res} \textbf{42}: D643--D648.

\leavevmode\hypertarget{ref-Zhu2019}{}%
Zhu, H. (2019) kableExtra: Construct complex table with 'kable' and pipe
syntax.

\newpage 
\setlength\parindent{0pt}

\hypertarget{figure-captions}{%
\subsection{Figure Captions}\label{figure-captions}}

\textbf{\autoref{confocal_december}.} \nameref{confocal_december}

\textbf{\autoref{confocal_june}.} \nameref{confocal_june}

\textbf{\autoref{community}.} \nameref{community}

\textbf{\autoref{chloroplast}.} \nameref{chloroplast}

\textbf{\autoref{cog}.} \nameref{cog}

\newpage

\hypertarget{figures}{%
\subsection{Figures}\label{figures}}

\begin{figure}[ht]

{\centering \includegraphics[width=0.85\linewidth]{../results/images/confocal_december}

}

\caption{Confocal microscope images of \textit{C. nodosa} and \textit{C. cylindracea} surfaces from the Bay of Saline and the Bay of Funtana (mixed and monospecific settlements) sampled on 4 December 2017 and stained with SYBR Green I. Scale bar at all images is 60 \si{\um}.\label{confocal_december}}\label{fig:unnamed-chunk-1}
\end{figure}

\newpage
\begin{figure}[ht]

{\centering \includegraphics[width=0.85\linewidth]{../results/images/confocal_june}

}

\caption{Confocal microscope images of \textit{C. nodosa} and \textit{C. cylindracea} surfaces from the Bay of Funtana (mixed and monospecific settlements) sampled on 19 June 2018 and stained with SYBR Green I. Scale bar at all images is 60 \si{\um}.\label{confocal_june}}\label{fig:unnamed-chunk-2}
\end{figure}

\newpage
\begin{figure}[ht]

{\centering \includegraphics[width=1\linewidth]{../results/figures/community_bar_plot}

}

\caption{Taxonomic classification and relative contribution of the most abundant ($\geq$ 1 \si{\percent}) bacterial and archaeal sequences from surfaces of two marine macrophytes (\textit{C. nodosa} and \textit{C. cylindracea}) sampled in the Bay of Saline and the Bay of Funtana (mixed and monospecific settlements) and in two contrasting seasons (4 December 2017 and 19 June 2018).\label{community}}\label{fig:unnamed-chunk-3}
\end{figure}

\newpage
\begin{figure}[ht]

{\centering \includegraphics[width=1\linewidth]{../results/figures/chloroplast}

}

\caption{Taxonomic classification and relative contribution of chloroplast sequences from surfaces of two marine macrophytes (\textit{C. nodosa} and \textit{C. cylindracea}) sampled in the Bay of Saline and the Bay of Funtana (mixed and monospecific settlements) and in two contrasting seasons (4 December 2017 and 19 June 2018).\label{chloroplast}}\label{fig:unnamed-chunk-4}
\end{figure}

\newpage
\begin{figure}[ht]

{\centering \includegraphics[width=1\linewidth]{../results/figures/cog}

}

\caption{Relative contribution of each COG category to the total number of annotated coding sequences (A) from metagenomes and identified proteins (B) from metaproteomes associated with surfaces of two marine macrophytes (\textit{C. nodosa} and \textit{C. cylindracea}) sampled in the Bay of Saline and the Bay of Funtana (mixed and monospecific settlements) and in two contrasting seasons (4/14 December 2017 and 19 June 2018). Total number of annotated coding sequences and identified proteins is given above the corresponding bar.\label{cog}}\label{fig:unnamed-chunk-5}
\end{figure}

\end{document}