### 2021-02-09 submission/manuscript.tex
Fromat text.

\documentclass[12pt,]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=1.0in]{geometry}
\usepackage{hyperref}
\hypersetup{unicode=true,
            pdftitle={Selective DNA and Protein Isolation from Marine Macrophyte Surfaces},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

%%% Use protect on footnotes to avoid problems with footnotes in titles
\let\rmarkdownfootnote\footnote%
\def\footnote{\protect\rmarkdownfootnote}

%%% Change title format to be more compact
\usepackage{titling}

% Create subtitle command for use in maketitle
\providecommand{\subtitle}[1]{
  \posttitle{
    \begin{center}\large#1\end{center}
    }
}

\setlength{\droptitle}{-2em}

  \title{\textbf{Selective DNA and Protein Isolation from Marine Macrophyte
Surfaces}}
    \pretitle{\vspace{\droptitle}\centering\huge}
  \posttitle{\par}
    \author{}
    \preauthor{}\postauthor{}
    \date{}
    \predate{}\postdate{}

\usepackage{times} % Times New Roman font
\usepackage[T1]{fontenc}

\usepackage[none]{hyphenat}

\usepackage{setspace}
\doublespacing
\setlength{\parskip}{1em}

\usepackage{lineno}
\renewcommand{\linenumberfont}{\normalfont\tiny}

\usepackage{pdfpages}

\usepackage{indentfirst}

\usepackage[labelsep=period, labelfont=bf]{caption}
\renewcommand{\thefigure}{\arabic{figure}}
\renewcommand{\figurename}{Fig.}
\captionsetup{justification=raggedright,singlelinecheck=false}

\usepackage{pdflscape}
\newcommand{\blandscape}{\begin{landscape}}
\newcommand{\elandscape}{\end{landscape}}

\usepackage{siunitx}
\DeclareSIUnit\molar{\mole\per\cubic\deci\metre}
\DeclareSIUnit\Molar{\textsc{m}}
\DeclareSIUnit\cells{\text{cells}}

\usepackage{caption}
\captionsetup{justification=justified}

\usepackage{float}

\usepackage{xr}
\externaldocument[supp-]{supplementary}

\usepackage{txfonts}

\renewcommand{\figureautorefname}{Fig.}

\usepackage{setspace}

\begin{document}
\maketitle

\vspace{20mm}

Marino Korlević\textsuperscript{1\(*\)}, Marsej
Markovski\textsuperscript{1}, Zihao Zhao\textsuperscript{2}, Gerhard J.
Herndl\textsuperscript{2,3,4}, Mirjana Najdek\textsuperscript{1}

1. Center for Marine Research, Ruđer Bošković Institute, Croatia

2. Department of Functional and Evolutionary Ecology, University of
Vienna, Austria

3. NIOZ, Department of Marine Microbiology and Biogeochemistry, Royal
Netherlands Institute for Sea Research, Utrecht University, The
Netherlands

4. Vienna Metabolomics Center, University of Vienna, Austria

\textsuperscript{\(*\)}To whom correspondence should be addressed:

Marino Korlević

G. Paliaga 5, 52210 Rovinj, Croatia

Tel.: +385 52 804 768

Fax: +385 52 804 780

e-mail:
\href{mailto:marino.korlevic@irb.hr}{\nolinkurl{marino.korlevic@irb.hr}}

Running title: DNA and protein isolation from macrophyte surfaces

\newpage
\linenumbers
\sisetup{mode=text}
\setlength\parindent{24pt}
\doublespacing

\hypertarget{summary}{%
\subsection{Summary}\label{summary}}

Studies of unculturable microbes often combine methods such as 16S rRNA
sequencing, metagenomics and metaproteomics. To apply these techniques
to the microbial community inhabiting the surfaces of marine macrophytes
it is advisable to perform a selective DNA and protein isolation prior
to the analysis to avoid biases due to the host material being present
in high quantities. Two protocols for DNA and protein isolation were
adapted for selective extractions of DNA and proteins from epiphytic
communities inhabiting the surfaces of two marine macrophytes, the
seagrass \emph{Cymodocea nodosa} and the macroalga \emph{Caulerpa
cylindracea}. Protocols showed an almost complete removal of the
epiphytic community regardless of the sampling season, station,
settlement or host species. The obtained DNA was suitable for
metagenomic and 16S rRNA sequencing, while isolated proteins could be
identified by mass spectrometry. Low presence of host DNA and proteins
in the samples indicated a high specificity of the protocols. The
procedures are based on universally available laboratory chemicals
making the protocols widely applicable. Taken together, the adapted
protocols ensure an almost complete removal of the macrophyte epiphytic
community. The procedures are selective for microbes inhabiting
macrophyte surfaces and provide DNA and proteins applicable in 16S rRNA
sequencing, metagenomics and metaproteomics.

\newpage

\hypertarget{introduction}{%
\subsection{Introduction}\label{introduction}}

Surfaces of marine macrophytes are colonized by a diverse microbial
community whose structure and function are poorly understood (Egan et
al., 2013). As less than 1 \% of all prokaryotic species are culturable,
molecular methods such as 16S rRNA sequencing, metagenomics and
metaproteomics are indispensable to study these organisms (Amann et al.,
1995; Su et al., 2012). Applying these techniques requires an initial
isolation step with the purpose of obtaining high quality DNA and
proteins.

Biological material (i.e., proteins and DNA) from pelagic microbial
communities is usually isolated by collecting cells onto filters and
subsequently isolating the target organisms or communities (Gilbert et
al., 2009). If a specific microbial size fraction is aimed sequential
filtration is applied (Massana et al., 1997; Andersson et al., 2010). In
contrast, obtaining microorganisms associated to surfaces require either
a cell detachment procedure prior to isolation or the host material is
co-extracted with the target material. Methods for separating microbial
cells from the host include shaking of host tissue (Gross et al., 2003;
Nõges et al., 2010), scraping of macrophyte surfaces (Uku et al., 2007)
or applying ultrasonication (Weidner et al., 1996; Cai et al., 2014). It
was shown that shaking alone is not sufficient to remove microbial cells
from surfaces, at least not from plant root surfaces (Richter-Heitmann
et al., 2016). Manual separation methods, such as scraping and brushing
are time consuming and subjective, as the detachment efficiency depends
on host tissue and the person performing the procedure (Cai et al.,
2014). Ultrasonication was proposed as an alternative method as it is
providing better results in terms of detachment efficiency (Cai et al.,
2014; Richter-Heitmann et al., 2016). The downside of this procedure is
that complete cell removal is still not obtained and tissue disruption
was observed especially after the application of probe ultrasonication
(Richter-Heitmann et al., 2016). An alternative to these cell detachment
procedures is the isolation of target epiphytic compounds together with
host material (Staufenberger et al., 2008; Jiang et al., 2015). This
procedure can lead to problems in the following processing steps such as
mitochondrial and chloroplast 16S rRNA sequence contaminations from the
host (Longford et al., 2007; Staufenberger et al., 2008). In addition,
when performing metagenomics and metaproteomics host material can cause
biased results towards more abundant host DNA and proteins.

An alternative to these procedures is a direct isolation of the target
material by incubating macrophyte tissues in an extraction buffer. After
the incubation, the undisrupted host tissue is removed followed by the
isolation procedure, omitting host material contaminations. To our
knowledge, the only procedure describing a direct and selective
epiphytic DNA isolation from the surfaces of marine macrophytes was
described by Burke et al. (2009). In contrast to previously described
methods, this protocol enables an almost complete removal of the surface
community. It was used for 16S rRNA gene clone library construction
(Burke et al., 2011b) and metagenome sequencing (Burke et al., 2011a).
This method, although providing a selective isolation procedure, uses a
rapid multi-enzyme cleaner (3M) that is not available worldwide and the
chemical constituents are unknown (Burke et al., 2009). Also to our
knowledge, no selective isolation protocol to perform (meta)proteomics
of epiphytic communities associated with marine macrophytes has been
developed yet.

In the present study, we adapted a protocol widely used for DNA
isolation from filters (Massana et al., 1997) and a protocol used for
protein isolation from soils (Chourey et al., 2010; Hultman et al.,
2015). These two adapted methods allowed for a selective extraction of
DNA and proteins from epiphytic communities inhabiting the surfaces of
two marine macrophytes, the seagrass \emph{Cymodocea nodosa} and the
macroalga \emph{Caulerpa cylindracea}. In addition, we tested the
removal efficiency of the protocol and the suitability of obtained DNA
and proteins for 16S rRNA sequencing, metagenomics and metaproteomics.

\newpage

\hypertarget{materials-and-methods}{%
\subsection{Materials and methods}\label{materials-and-methods}}

\hypertarget{sampling}{%
\subsubsection{Sampling}\label{sampling}}

Leaves of \emph{C. nodosa} were sampled in a \emph{C. nodosa} meadow in
the Bay of Saline, northern Adriatic Sea (\ang{45;7;5} N, \ang{13;37;20}
E) and in a \emph{C. nodosa} meadow invaded by \emph{C. cylindracea} in
the Bay of Funtana, northern Adriatic Sea (\ang{45;10;39} N,
\ang{13;35;42} E). Thalli of \emph{C. cylindracea} were sampled in the
same \emph{C. nodosa} invaded meadow in the Bay of Funtana and at a
locality of only \emph{C. cylindracea} located in the proximity of the
invaded meadow. Leaves and thalli for 16S rRNA analysis, metagenomics
and metaproteomics were collected in two contrasting seasons, on 4
December 2017 (16S rRNA analysis and metaproteomics), 14 December 2017
(metagenomics) and 18 June 2018 (16S rRNA analysis, metagenomics and
metaproteomics). During spring 2018, the \emph{C. nodosa} meadow in the
Bay of Saline decayed to an extent that no leaves could be retrieved
(Najdek et al., 2020). In addition, as not enough DNA for both
metagenomic and 16S RNA analysis were obtained during the sampling on 4
December 2017, an additional sampling on 14 December 2017 was carried
out in the Bay of Funtana. Leaves and thalli were collected by diving
and transported to the laboratory in containers placed on ice and filled
with seawater from this site. Upon arrival to the laboratory, \emph{C.
nodosa} leaves were cut into sections of 1 -- 2 \si{\cm}, while \emph{C.
cylindracea} thalli were cut into 5 -- 8 \si{\cm} long sections. Leaves
and thalli were washed three times with sterile artificial seawater
(ASW) to remove loosely attached microbial cells.

\hypertarget{dna-isolation}{%
\subsubsection{DNA isolation}\label{dna-isolation}}

The DNA was isolated according to the protocol for isolation from
filters described in Massana et al. (1997). This protocol was modified
and adapted for microbial DNA isolation from macrophyte surfaces as
described below. Five (5) \si{\ml} of lysis buffer (40 \si{\milli\Molar}
EDTA, 50 \si{\milli\Molar} Tris-HCl, 0.75 \si{\Molar} sucrose; pH 8.3)
was added to 1 \si{\g} wet weight of leaves or 2 \si{\g} wet-weight of
thalli. For every sample, duplicate isolations were performed. Lysozyme
was added (final concentration 1 \si{\mg\per\ml}) and the mixture was
incubated at 37 \si{\degreeCelsius} for 30 \si{\minute}. Subsequently,
proteinase K (final concentration 0.5 \si{\mg\per\ml}) and SDS (final
concentration 1 \si{\percent}) were added and the samples were incubated
at 55 \si{\degreeCelsius} for 2 \si{\hour}. Following the incubation,
tubes were vortexed for 10 \si{\minute} and the mixture containing lysed
epiphytic cells was separated from host leaves or thalli by transferring
the solution into a clean tube. The lysate was extracted twice with a
mixture of phenol:chloroform:isoamyl alcohol (25:24:1; pH 8) and once
with chloroform:isoamyl alcohol (24:1). After each addition of an
organic solvent mixture, tubes were slightly vortexed and centrifuged at
4,500 × g for 10 \si{\minute}. Following each centrifugation the aqueous
phases were retrieved. After the final extraction 1/10 of chilled 3
\si{\Molar} sodium acetate (pH 5.2) was added. DNA was precipitated by
adding 1 volume of chilled isopropanol, incubating the mixtures
overnight at \num{-20} \si{\degreeCelsius} and centrifuging at 16,000 ×
g and 4 \si{\degreeCelsius} for 20 \si{\minute}. The pellet was washed
twice with 1 \si{\ml} of chilled 70 \si{\percent} ethanol and
centrifuged after each washing step at 20,000 × g and 4
\si{\degreeCelsius} for 10 \si{\minute}. After the first washing step
duplicate pellets from the same sample were pooled and transferred to a
clean 1.5 \si{\ml} tube. The dried pellet was re-suspended in 100
\si{\ul} of deionized water.

\hypertarget{illumina-16s-rrna-sequencing}{%
\subsubsection{Illumina 16S rRNA
sequencing}\label{illumina-16s-rrna-sequencing}}

An aliquot of isolated DNA was treated with RNase A (final concentration
200 \si{\ug\per\ml}) for 2 \si{\hour} at 37 \si{\degreeCelsius}. The DNA
concentration was determined using the Quant-iT PicoGreen dsDNA Assay
Kit (Invitrogen, USA) according to the manufacturer's instructions and
diluted to 1 \si{\ng\per\ul}. The V4 region of the 16S rRNA gene was
amplified using a two-step PCR procedure. In the first PCR the 515F
(\(5'\)-GTGYCAGCMGCCGCGGTAA-\(3'\)) and 806R
(\(5'\)-GGACTACNVGGGTWTCTAAT-\(3'\)) primers from the Earth Microbiome
Project
(\url{http://press.igsb.anl.gov/earthmicrobiome/protocols-and-standards/16s/})
were used to amplify the target region (Caporaso et al., 2012; Apprill
et al., 2015; Parada et al., 2016). These primers contained on their
\(5'\) end a tagged sequence. Each sample was amplified in four parallel
25 \si{\ul} reactions of which each contained 1 × Q5 Reaction Buffer,
0.2 \si{\milli\Molar} of dNTPmix, 0.7 \si{\mg\per\ml} BSA (Bovine Serum
Albumin), 0.2 \si{\micro\Molar} of forward and reverse primers, 0.5 U of
Q5 High-Fidelity DNA Polymerase (New England Biolabs, USA) and 5
\si{\ng} of DNA template. Cycling conditions were: initial denaturation
at 94 \si{\degreeCelsius} for 3 \si{\minute}, 20 cycles of denaturation
at 94 \si{\degreeCelsius} for 45 \si{\s}, annealing at 50
\si{\degreeCelsius} for 60 \si{\s} and elongation at 72
\si{\degreeCelsius} for 90 \si{\s}, finalized by an elongation step at
72 \si{\degreeCelsius} for 10 \si{\minute}. The four parallel reactions
volumes were pooled and PCR products were purified using the GeneJET PCR
Purification Kit (Thermo Fisher Scientific, USA) according to the
manufacturer's instructions and following the protocol that included
isopropanol addition for better small DNA fragment yield. The column was
eluted in 30 \si{\ul} of deionized water. Purified PCR products were
sent for Illumina MiSeq sequencing (2 × 250 bp) at IMGM Laboratories,
Martinsried, Germany. Before sequencing at IMGM, the second PCR
amplification of the two-step PCR procedure was performed using primers
targeting the tagged region incorporated in the first PCR. In addition,
these primers contained adapter and sample-specific index sequences. The
second PCR was carried out for 8 cycles. Beside samples, a positive and
negative control were sequenced. A negative control was comprised of
four parallel PCR reactions without DNA template, while for a positive
control a mock community composed of evenly mixed DNA material
originating from 20 bacterial strains (ATCC MSA-1002, ATCC, USA) was
used. Partial 16S rRNA sequences obtained in this study have been
deposited in the European Nucleotide Archive (ENA) at EMBL-EBI under
accession numbers SAMEA6786270, SAMEA6648792 -- SAMEA6648794,
SAMEA6648809 -- SAMEA6648811 and SAMEA6648824.

Obtained sequences were analysed on the computer cluster Isabella
(University Computing Center, University of Zagreb) using mothur
(version 1.43.0) (Schloss et al., 2009) according to the MiSeq Standard
Operating Procedure (MiSeq SOP; \url{https://mothur.org/wiki/MiSeq_SOP})
(Kozich et al., 2013) and recommendations given from the Riffomonas
project to enhance data reproducibility
(\url{http://www.riffomonas.org/}). For alignment and classification of
sequences the SILVA SSU Ref NR 99 database (release 138;
\url{http://www.arb-silva.de}) was used (Quast et al., 2013; Yilmaz et
al., 2014). Sequences classified as chloroplasts by SILVA were exported
and reclassified using mothur and the RDP
(\url{http://rdp.cme.msu.edu/}) training set (version 16) reference
files adapted for mothur (Cole et al., 2014). In comparison to SILVA,
RDP allows a more detailed classification of chloroplast sequences.
Based on the ATCC MSA-1002 mock community included in the analysis a
sequencing error rate of 0.009 \si{\percent} was determined, which is in
line with previously reported values for next-generation sequencing data
(Kozich et al., 2013; Schloss et al., 2016). In addition, the negative
control processed together with the samples yielded only 2 sequences
after sequence quality curation.

\hypertarget{metagenomics}{%
\subsubsection{Metagenomics}\label{metagenomics}}

Four DNA samples were selected and sent on dry ice to IMGM Laboratories,
Martinsried, Germany for metagenomic sequencing. DNA was purified using
AMPure XP Beads (Beckman Coulter, USA) applying a bead:DNA ratio of 1:1
(v/v), quantified with a Qubit dsDNA HS Assay Kit (Thermo Fisher
Scientific, USA) and check for integrity on a 1 \si{\percent} agarose
gel. Metagenomic sequencing libraries were generated from 300 \si{\ng}
of input DNA using a NEBNext Ultra II FS DNA Library Prep Kit for
Illumina (New England Biolabs, USA) according to the manufacturer's
instructions. Fragments were selected (500 -- 700 bp) using AMPure XP
Beads, PCR enriched for 3 -- 5 cycles and quality controlled. Libraries
generated from different DNA samples were pooled and sequenced on an
Illumina NovaSeq 6000 sequencing system (2 × 250 bp).

Obtained sequences were analysed on the Life Science Compute Cluster
(LiSC) (CUBE -- Computational Systems Biology, University of Vienna).
Individual sequences were assembled using MEGAHIT (version 1.1.2) (Li et
al., 2015) under default settings. Putative genes were predicted from
contigs longer than 200 bp using Prodigal (version 2.6.3) (Hyatt et al.,
2010) in metagenome mode (-p meta). Abundances of predicted genes were
expressed as Reads Per Kilobase Million (RPKM) and calculated using the
BWA algorithm (version 0.7.16a) (Li and Durbin, 2009). All predicted
genes were functionally annotated using the eggNOG-mapper (Huerta-Cepas
et al., 2017) and eggNOG database (version 5.0) (Huerta-Cepas et al.,
2019). Sequence taxonomy classification was determined using the lowest
common ancestor algorithm adapted from DIAMOND (version 0.8.36)
(Buchfink et al., 2015) and by searching against the NCBI non-redundant
database (NR). To determine the phylogeny, the top 10 \si{\percent} hits
with an e-value \textless{} 1 × 10\textsuperscript{-5} were used
(-\/-top 10). Sequence renaming, coverage information computing and
metagenomic statistics calculations were performed using software tools
from BBTools (\url{https://jgi.doe.gov/data-and-tools/bbtools}).
Metagenomic sequences obtained in this study have been deposited in the
European Nucleotide Archive (ENA) at EMBL-EBI under accession numbers
SAMEA6648795, SAMEA6648797, SAMEA6648809 and SAMEA6648811.

\hypertarget{protein-isolation}{%
\subsubsection{Protein isolation}\label{protein-isolation}}

Proteins were isolated according to the protocol for protein isolation
from soil described in Chourey et al. (2010) and modified by Hultman et
al. (2015). This protocol was further modified and adapted for microbial
protein isolation from macrophyte surfaces as described below. Twenty
(20) \si{\ml} of protein extraction buffer (4 \si{\percent} SDS, 100
\si{\milli\Molar} Tris-HCl {[}pH 8.0{]}) was added to 5 \si{\g} wet
weight of leaves or 10 \si{\g} wet weight of thalli. The mixture was
incubated in boiling water for 5 \si{\minute}, vortexed for 10
\si{\minute} and incubated again in boiling water for 5 \si{\minute}.
After a brief vortex, the lysate was transferred to a clean tube
separating the host leaves or thalli from the mixture containing lysed
epiphytic cells. Dithiothreitol (DTT; final concentration 24
\si{\milli\Molar}) was added and proteins were precipitated with chilled
100 \si{\percent} trichloroacetic acid (TCA; final concentration 20
\si{\percent}) overnight at \num{-20} \si{\degreeCelsius}. Precipitated
proteins were centrifuged at 10,000 × g and 4 \si{\degreeCelsius} for 40
\si{\minute}. The obtained protein pellet was washed three times with
chilled acetone. During the first washing step the pellet was
transferred to a clean 1.5 \si{\ml} tube. After each washing step,
samples were centrifuged at 20,000 × g and 4 \si{\degreeCelsius} for 5
\si{\minute}. Dried pellets were stored at \num{-80} \si{\degreeCelsius}
until further analysis.

\hypertarget{metaproteomics}{%
\subsubsection{Metaproteomics}\label{metaproteomics}}

Isolated proteins were whole trypsin digested using the FASP
(filter-aided sample preparation) Protein Digestion Kit (Expedeon, UK)
according to the manufacturer's instructions with small modifications
(Wiśniewski et al., 2009). Prior to loading the solution onto the
column, protein pellets were solubilized in a urea sample buffer
included in the kit amended with DTT (final concentration 100
\si{\milli\Molar}) for 45 \si{\minute} at room temperature and
centrifuged at 20,000 × g for 2 -- 5 \si{\minute} at room temperature to
remove larger particles. The first washing step after protein solution
loading was repeated twice. In addition, the centrifugation steps were
prolonged if the column was clogged. Trypsin digestion was performed on
column filters at 37 \si{\degreeCelsius} overnight for 18 \si{\hour}.
The final filtrate containing peptides was acidified with 1
\si{\percent} (final concentration) trifluoroacetic acid (TFA), freezed
at \num{-80} \si{\degreeCelsius}, lyophilized and sent to VIME -- Vienna
Metabolomics Center (University of Vienna) for metaproteomic analysis.
Peptides were re-suspended in 1 \si{\percent} (final concentration) TFA,
desalted using the Pierce C18 Tips (Thermo Fisher Scientific, USA)
according to the manufacturer's instructions and sequenced on a Q
Exactive Hybrid Quadrupole-Orbitrap Mass Spectrometer (Thermo Fisher
Scientific, USA). Obtained MS/MS spectra were searched against a protein
database composed of combined sequenced metagenomes using SEQUEST-HT
engines and validated with Percolator in Proteome Discoverer 2.1 (Thermo
Fisher Scientific, USA). The target-decoy approach was used to reduce
the probability of false peptide identification. Results whose false
discovery rate at the peptide level was \textless{} \SI{1}{\percent}
were kept. For protein identification a minimum of two peptides and one
unique peptide were required. For protein quantification, a
chromatographic peak area-based free quantitative method was applied.

\hypertarget{data-processing-and-visualization}{%
\subsubsection{Data processing and
visualization}\label{data-processing-and-visualization}}

Processing and visualization of 16S rRNA, metagenomic and metaproteomic
data were done using R (version 3.6.0) (R Core Team, 2019), package
tidyverse (version 1.3.0) (Wickham et al., 2019) and multiple other
packages (Neuwirth, 2014; Xie, 2014, 2015, 2019, 2020; Wilke, 2018; Xie
et al., 2018; Allaire et al., 2019; Zhu, 2019; Bengtsson, 2020). The
detailed analysis procedure including the R Markdown file for this paper
are available as a GitHub repository
(\url{https://github.com/MicrobesRovinj/Korlevic_SelectiveRemoval_FrontMicrobiol_2021}).

\hypertarget{confocal-microscopy}{%
\subsubsection{Confocal microscopy}\label{confocal-microscopy}}

Host leaves and thalli from DNA and protein isolation steps were washed
seven times in deionized water and fixed with formaldehyde (final
concentration \textasciitilde{}3 \si{\percent}). In addition,
non-treated leaves and thalli, washed three times in ASW to remove
loosely attached microbial cells, were fixed in the same concentration
of formaldehyde and used as a positive control. For-long term storage,
fixed leaves and thalli were immersed in a mixture of phosphate-buffered
saline (PBS) and ethanol (1:1) and stored at \num{-20}
\si{\degreeCelsius}. Treated and untreated segments of leaves and thalli
were stained in a 2 × solution of SYBR Green I and examined under a
Leica TCS SP8 X FLIM confocal microscope (Leica Microsystems, Germany).

\newpage

\hypertarget{results}{%
\subsection{Results}\label{results}}

To assess the removal efficiency of the DNA and protein isolation
procedures, leaves and thalli were examined under a confocal microscope
before and after treatments were performed. The modified procedures
resulted in an almost complete removal of the surface community of both,
\emph{C. nodosa} and \emph{C. cylindracea}. In addition, a similar
removal efficiency was observed for communities sampled in contrasting
months, December 2017 (\autoref{confocal_december}) and June 2018
(\autoref{confocal_june}). Also, no effect of station, settlement or
isolation procedure (DNA or protein) on the removal efficiency was
observed (Figs. \ref{confocal_december} and \ref{confocal_june}).

To evaluate whether the obtained DNA is suitable to determine the
composition of the microbial community Illumina sequencing of the V4
region of the 16S rRNA was performed. Sequencing yielded a total of
336,944 sequences after quality curation and exclusion of eukaryotic,
mitochondrial and no relative sequences. The number of sequences
classified as chloroplasts was 97,334. After excluding these sequences,
the total number of retrieved reads was 239,610, ranging from 22,596 to
52,930 sequences per sample (\autoref{supp-nseq_notus}). Even when the
highest sequencing effort was applied, the rarefaction curves did not
level off which is commonly observed in high-throughput 16S rRNA
amplicon sequencing (\autoref{supp-rarefaction}). Sequences clustering
at a similarity level of 97 \si{\percent} yielded a total of 8,355
different OTUs. Taxonomic classification of reads revealed a
macrophyte-associated epiphytic community that mainly composed of
\emph{Alphaproteobacteria} (14.9 ± 3.5 \si{\percent}),
\emph{Bacteroidota} (12.5 ± 2.4 \si{\percent}),
\emph{Gammaproteobacteria} (11.6 ± 4.3 \si{\percent}),
\emph{Desulfobacterota} (7.8 ± 7.5 \si{\percent}), \emph{Cyanobacteria}
(6.5 ± 4.7 \si{\percent}) and \emph{Planctomycetota} (2.9 ± 1.7
\si{\percent}) (\autoref{community}).

Primers used in this study, as in many other 16S rRNA amplicon
procedures, also amplified chloroplast 16S rRNA genes. We observed a
high proportion of chloroplast sequences in all analysed samples (33.4 ±
9.4 \si{\percent}) (\autoref{community}). To determine whether
chloroplast sequences originate from the host or eukaryotic epiphytic
organisms, we exported SILVA-classified chloroplast sequences and
reclassified them using the RDP (Ribosomal Database Project) training
set that allows for a more detailed chloroplast classification. The
largest proportion of sequences was classified as Bacillariophyta (89.7
± 5.7 \si{\percent}) indicating that the DNA removal procedure resulted
in only minor co-extracted quantities of host DNA
(\autoref{chloroplast}). Chloroplast sequences classified as
Streptophyta constituted 3.3 ± 2.8 \si{\percent} of all chloroplast
sequences originating from \emph{C. nodosa} samples, while sequences
classified as Chlorophyta comprised only 0.02 ± 0.01 \si{\percent} of
all chloroplast sequences associated with \emph{C. cylindracea} samples.

To determine whether the extracted DNA can be used for metagenomic
sequencing, four samples containing epiphytic DNA were selected and
shotgun sequenced using an Illumina platform. Metagenomic sequencing
yielded between 207,149,524 and 624,029,930 sequence pairs
(\autoref{supp-metagenomic_statistics}). Obtained sequences were
successfully assembled into contigs whose L50 ranged from 654 to 1,011
bp. In addition, predicted coding sequences were functionally annotated
(9,066,667 -- 20,256,215 annotated sequences; \autoref{cog}a) and
taxonomically classified. Functional annotation allowed for an
assessment of the relative contribution of each COG (Clusters of
Orthologous Groups) functional category to the total number of annotated
coding sequences (\autoref{cog}a). Functional categories containing the
highest number of sequences were C (Energy production and conversion), E
(Amino acid transport and metabolism), M (Cell wall/membrane/envelope
biogenesis), L (Replication, recombination and repair) and P (Inorganic
ion transport and metabolism). If host DNA is co-extracted with
epiphytes it should be detected in large proportions in sequenced
metagenomes. However, no large proportions of coding sequences
classified as Streptophyta and Chlorophyta were detected
(\autoref{supp-metagenomic_taxonomy}). Sequenced metagenomic DNA
originating from the surface of \emph{C. nodosa} contained 1.3
\si{\percent} of coding sequences classified as Streptophyta in December
2017 and 0.7 \si{\percent} in June 2018. Furthermore, the summed RPKM
(Reads Per Kilobase Million) of these sequences constituted 1.7
\si{\percent} of total RPKM of all successfully classified sequences in
December 2017 and 1.1 \si{\percent} in June 2018. Similar low
proportions of host coding sequences were detected in metagenomic
samples originating from the surfaces of \emph{C. cylindracea}. Of all
successfully classified coding sequences 0.2 \si{\percent} were
classified as Chlorophyta in December 2017 and 0.1 \si{\percent} in June
2018. A relatively higher proportion of RPKM of these sequences than in
the case of \emph{C. nodosa} was observed, indicating a higher
co-extraction of host DNA in \emph{C. cylindracea}. In December, the
proportion of RPKM of sequences classified as Chlorophyta was 8.2
\si{\percent}, while in June 2018 it reached 13.6 \si{\percent}.

To evaluate whether the procedure for protein extraction is suitable for
metaproteomic analysis, obtained proteins were trypsin digested and
sequenced using a mass spectrometer. Obtained MS/MS spectra were
searched against a protein database from sequenced metagenomes. From
14,219 to 16,449 proteins were identified in isolated protein samples
(\autoref{cog}b). In addition, successful identification of proteins
allowed for an assessment of the relative contribution of each COG
functional category to the total number of identified proteins
(\autoref{cog}b). Functional categories containing the highest number of
identified proteins were C (Energy production and conversion), G
(Carbohydrate transport and metabolism), P (Inorganic ion transport and
metabolism), O (Posttranslational modification, protein turnover,
chaperones) and E (Amino acid transport and metabolism). Isolated
proteins could originate from epiphytic organisms inhabiting the
macrophyte surface and/or from macrophyte tissue underlying them. The
contribution of proteins originating from host tissues was evaluated by
identifying all the proteins predicted to belong to any taxonomic group
within the phyla Streptophyta and Chlorophyta and by calculating their
contribution to the number and abundance (NAAF -- Normalized Abundance
Area Factor) of all identified proteins. On average, proteins isolated
from the surface of \emph{C. nodosa} contained 1.8 ± 0.06 \si{\percent}
of proteins associated with Streptophyta, contributing to 2.2 ± 0.8
\si{\percent} of total proteins. Similar to metagenomes, proteins
associated with Chlorophyta contributed more to \emph{C. cylindracea}
than proteins associated with Streptophyta to \emph{C. nodosa}.
Chlorophyta associated proteins comprised 5.2 ± 0.06 \si{\percent} of
all identified proteins in \emph{C. cylindracea}, contributing 19.2 ±
1.5 \si{\percent} to the total protein abundance.

\newpage

\hypertarget{discussion}{%
\subsection{Discussion}\label{discussion}}

To test whether the developed DNA and protein isolation protocols
efficiently detach microbes from the macrophyte surface we selected
\emph{C. nodosa} and \emph{C. cylindracea} as representatives of
seagrass and macroalgal species. These species differ morphologically.
While \emph{C. nodosa} leaves are flat, \emph{C. cylindracea} thalli are
characterized by an uneven surface (Kuo and den Hartog, 2001; Verlaque
et al., 2003). The developed protocol led to an almost complete removal
of epiphytic cells from the surfaces of both species comparable to the
result of Burke et al. (2009), indicating that structural differences do
not impact the removal efficiency. In addition, isolation protocols were
tested in two contrasting seasons, as it is known that macrophytes are
harbouring more algal epiphytes during autumn and winter (Reyes and
Sansón, 2001). No differences in the removal efficiency was observed
between seasons suggesting that these protocols can be used on
macrophyte samples retrieved throughout the year. Also, no removal
differences were observed on samples derived from the same host but from
different locations.

Successful amplification and sequencing of the V4 region of the 16S rRNA
gene proved that the isolated DNA can be used to estimate the microbial
epiphytic diversity. Taxonomic groups detected in this step can also be
often found in epiphytic communities associated with other macrophytes
(Burke et al., 2011b; Morrissey et al., 2019). A problem often
encountered in studies focusing on epiphytic communities is the presence
of large proportions of chloroplast 16S rRNA sequences in the pool of
amplified molecules, especially if the epiphytic DNA was isolated
without prior selection (Staufenberger et al., 2008). These sequences
can derive from host chloroplasts or from eukaryotic epiphytic
chloroplast DNA. Although the proportion of obtained chloroplast 16S
rRNA sequences in our samples was substantial, they derived almost
exclusively from eukaryotic epiphytes. High proportion of chloroplast
16S rRNA sequences in studies applying selective procedures that include
direct cellular lysis on host surfaces were observed before (Michelou et
al., 2013). It is possible that chloroplast-specific sequences even in
these studies originated from eukaryotic epiphytic cells and not from
host chloroplasts. Indeed, it is common during 16S rRNA profiling of
pelagic microbial communities to observe high proportions of chloroplast
sequences (Gilbert et al., 2009; Korlević et al., 2016). In addition, a
very low proportion of chloroplast 16S rRNA sequences in samples
originating from \emph{C. cylindracea} in comparison to \emph{C. nodosa}
could be explained by the presence of three introns in the gene for 16S
rRNA in some members of the genus \emph{Caulerpa} that could hamper the
amplification process (Lam and Lopez-Bautista, 2016).

High quality DNA is also needed for metagenomics. The obtained number of
metagenomic sequences and assembly statistics were comparable to
metagenomes and metatranscriptomes derived from similar surface
associated communities (Crump et al., 2018; Cúcio et al., 2018). In
addition, functional annotation of predicted coding sequences to COG
functional categories showed that the obtained metagenomes can be used
to determine the metabolic capacity of surface associated communities
(Leary et al., 2014; Cúcio et al., 2018). The proportion of coding
sequences, including their RPKM, originating from \emph{C. nodosa}
metagenomes and classified as Streptophyta was low indicating that the
isolation procedure was specific for epiphytic cells. DNA samples
isolated from the surface of \emph{C. cylindracea} exhibited a low
proportion of Chlorophyta coding sequences, however, their RPKM was
higher than in the samples originating from \emph{C. nodosa}. One of the
reasons for this elevated RPKM of Chlorophyta sequences in \emph{C.
cylindracea} could be the differences in the tissue structure between
these two host species. While \emph{C. nodosa} leaves are composed of
individual cells, the thallus of \emph{C. cylindracea} is, like in other
siphonous algal species, composed of a single large multinucleate cell
(Coneva and Chitwood, 2015). The absence of individual cells in \emph{C.
cylindracea} could cause a leakage of genetic material into the
extraction buffer causing an elevated presence of host sequences in the
samples for metagenome analyses.

To obtain insight into the metabolic status of uncultivated prokaryotes,
a metaproteomic approach is required (Saito et al., 2019). The applied
protocol for epiphytic protein isolation followed by a metaproteomic
analysis identified between 14,219 and 16,449 proteins, which is higher
than previously reported for e.g.~soils (Chourey et al., 2010; Hultman
et al., 2015), seawater (Williams et al., 2012) and biofilms (Leary et
al., 2014). The functional annotation of identified proteins into COG
functional categories showed that the protein isolation protocol can be
used to assess the metabolic status of the epiphytic community (Leary et
al., 2014). Similar to the results of the metagenomic analysis, the
number and abundance of identified proteins affiliated to Streptophyta
in \emph{C. nodosa} samples were low, indicating that the procedure is
selective for epiphytic cell proteins. In addition, a higher number and
abundance of identified proteins associated with Chlorophyta were
observed in \emph{C. cylindracea} samples. The cause of this elevated
presence of Chlorophyta-associated proteins can be, similar to the DNA
isolation protocol, explained by the absence of individual cells in this
siphonous alga (Coneva and Chitwood, 2015).

In conclusion, the developed protocols for DNA and protein isolation
from macrophyte surfaces almost completely remove the epiphytic
community from both, \emph{C. nodosa} and \emph{C. cylindracea}, in
different seasons. Also, the obtained DNA and proteins are suitable for
16S rRNA sequencing, metagenomics and metaproteomics analyses while the
obtained material contains low quantities of host DNA and proteins
making the protocols specific for epiphytes. Furthermore, the protocols
are based on universally available laboratory chemicals hence, making
them widely applicable.

\hypertarget{acknowledgements}{%
\subsection{Acknowledgements}\label{acknowledgements}}

This work was funded by the Croatian Science Foundation through the
MICRO-SEAGRASS project (project number IP-2016-06-7118). ZZ and GJH were
supported by the Austrian Science Fund (FWF) project ARTEMIS (project
number P28781-B21). We would like to thank Margareta Buterer for
technical support, Paolo Paliaga for help during sampling, Lucija Horvat
for technical support in confocal microscopy, Dušica Vujaklija for help
in accessing the confocal microscope and Sonja Tischler for technical
support in mass spectrometry. In addition, we would like to thank the
University Computing Center of the University of Zagreb for access to
the computer cluster Isabella and CUBE -- Computational Systems Biology
of the University of Vienna for access to the Life Science Compute
Cluster (LiSC).

\newpage

\hypertarget{references}{%
\subsection{References}\label{references}}

\interlinepenalty=10000 \setlength{\emergencystretch}{3.5em}

\hypertarget{refs}{}
\leavevmode\hypertarget{ref-Allaire2019}{}%
Allaire, J. J., Xie, Y., McPherson, J., Luraschi, J., Ushey, K., Atkins,
A., et al. (2019). \emph{rmarkdown: Dynamic documents for R}.

\leavevmode\hypertarget{ref-Amann1995}{}%
Amann, R. I., Ludwig, W., and Schleifer, K. H. (1995). Phylogenetic
identification and \emph{in situ} detection of individual microbial
cells without cultivation. \emph{Microbiol. Rev.} 59, 143--169.

\leavevmode\hypertarget{ref-Andersson2010}{}%
Andersson, A. F., Riemann, L., and Bertilsson, S. (2010). Pyrosequencing
reveals contrasting seasonal dynamics of taxa within Baltic Sea
bacterioplankton communities. \emph{ISME J.} 4, 171--181.
doi:\href{https://doi.org/10.1038/ismej.2009.108}{10.1038/ismej.2009.108}.

\leavevmode\hypertarget{ref-Apprill2015}{}%
Apprill, A., McNally, S., Parsons, R., and Weber, L. (2015). Minor
revision to V4 region SSU rRNA 806R gene primer greatly increases
detection of SAR11 bacterioplankton. \emph{Aquat. Microb. Ecol.} 75,
129--137. doi:\href{https://doi.org/10.3354/ame01753}{10.3354/ame01753}.

\leavevmode\hypertarget{ref-Bengtsson2020}{}%
Bengtsson, H. (2020). \emph{matrixStats: Functions that apply to rows
and columns of matrices (and to vectors)}.

\leavevmode\hypertarget{ref-Buchfink2015}{}%
Buchfink, B., Xie, C., and Huson, D. H. (2015). Fast and sensitive
protein alignment using DIAMOND. \emph{Nat. Methods} 12, 59--60.
doi:\href{https://doi.org/10.1038/nmeth.3176}{10.1038/nmeth.3176}.

\leavevmode\hypertarget{ref-Burke2009}{}%
Burke, C., Kjelleberg, S., and Thomas, T. (2009). Selective extraction
of bacterial DNA from the surfaces of macroalgae. \emph{Appl. Environ.
Microbiol.} 75, 252--256.
doi:\href{https://doi.org/10.1128/AEM.01630-08}{10.1128/AEM.01630-08}.

\leavevmode\hypertarget{ref-Burke2011}{}%
Burke, C., Steinberg, P., Rusch, D., Kjelleberg, S., and Thomas, T.
(2011a). Bacterial community assembly based on functional genes rather
than species. \emph{Proc. Natl. Acad. Sci. U. S. A.} 108, 14288--14293.
doi:\href{https://doi.org/10.1073/pnas.1101591108}{10.1073/pnas.1101591108}.

\leavevmode\hypertarget{ref-Burke2011a}{}%
Burke, C., Thomas, T., Lewis, M., Steinberg, P., and Kjelleberg, S.
(2011b). Composition, uniqueness and variability of the epiphytic
bacterial community of the green alga \emph{Ulva australis}. \emph{ISME
J.} 5, 590--600.
doi:\href{https://doi.org/10.1038/ismej.2010.164}{10.1038/ismej.2010.164}.

\leavevmode\hypertarget{ref-Cai2014}{}%
Cai, X., Gao, G., Yang, J., Tang, X., Dai, J., Chen, D., et al. (2014).
An ultrasonic method for separation of epiphytic microbes from
freshwater submerged macrophytes. \emph{J. Basic Microbiol.} 54,
758--761.
doi:\href{https://doi.org/10.1002/jobm.201300041}{10.1002/jobm.201300041}.

\leavevmode\hypertarget{ref-Caporaso2012}{}%
Caporaso, J. G., Lauber, C. L., Walters, W. A., Berg-Lyons, D., Huntley,
J., Fierer, N., et al. (2012). Ultra-high-throughput microbial community
analysis on the Illumina HiSeq and MiSeq platforms. \emph{ISME J.} 6,
1621--1624.
doi:\href{https://doi.org/10.1038/ismej.2012.8}{10.1038/ismej.2012.8}.

\leavevmode\hypertarget{ref-Chourey2010}{}%
Chourey, K., Jansson, J., VerBerkmoes, N., Shah, M., Chavarria, K. L.,
Tom, L. M., et al. (2010). Direct cellular lysis/protein extraction
protocol for soil metaproteomics. \emph{J. Proteome Res.} 9, 6615--6622.
doi:\href{https://doi.org/10.1021/pr100787q}{10.1021/pr100787q}.

\leavevmode\hypertarget{ref-Cole2014}{}%
Cole, J. R., Wang, Q., Fish, J. A., Chai, B., McGarrell, D. M., Sun, Y.,
et al. (2014). Ribosomal Database Project: Data and tools for high
throughput rRNA analysis. \emph{Nucleic Acids Res} 42, D633--D642.
doi:\href{https://doi.org/10.1093/nar/gkt1244}{10.1093/nar/gkt1244}.

\leavevmode\hypertarget{ref-Coneva2015}{}%
Coneva, V., and Chitwood, D. H. (2015). Plant architecture without
multicellularity: Quandaries over patterning and the soma-germline
divide in siphonous algae. \emph{Front. Plant Sci.} 6, 287.
doi:\href{https://doi.org/10.3389/fpls.2015.00287}{10.3389/fpls.2015.00287}.

\leavevmode\hypertarget{ref-Crump2018}{}%
Crump, B. C., Wojahn, J. M., Tomas, F., and Mueller, R. S. (2018).
Metatranscriptomics and amplicon sequencing reveal mutualisms in
seagrass microbiomes. \emph{Front. Microbiol.} 9, 388.
doi:\href{https://doi.org/10.3389/fmicb.2018.00388}{10.3389/fmicb.2018.00388}.

\leavevmode\hypertarget{ref-Cucio2018}{}%
Cúcio, C., Overmars, L., Engelen, A. H., and Muyzer, G. (2018).
Metagenomic analysis shows the presence of bacteria related to
free-living forms of sulfur-oxidizing chemolithoautotrophic symbionts in
the rhizosphere of the seagrass \emph{Zostera marina}. \emph{Front. Mar.
Sci.} 5, 171.
doi:\href{https://doi.org/10.3389/fmars.2018.00171}{10.3389/fmars.2018.00171}.

\leavevmode\hypertarget{ref-Egan2013}{}%
Egan, S., Harder, T., Burke, C., Steinberg, P., Kjelleberg, S., and
Thomas, T. (2013). The seaweed holobiont: Understanding seaweed-bacteria
interactions. \emph{FEMS Microbiol. Rev.} 37, 462--476.
doi:\href{https://doi.org/10.1111/1574-6976.12011}{10.1111/1574-6976.12011}.

\leavevmode\hypertarget{ref-Gilbert2009}{}%
Gilbert, J. A., Field, D., Swift, P., Newbold, L., Oliver, A., Smyth,
T., et al. (2009). The seasonal structure of microbial communities in
the Western English Channel. \emph{Environ. Microbiol.} 11, 3132--3139.
doi:\href{https://doi.org/10.1111/j.1462-2920.2009.02017.x}{10.1111/j.1462-2920.2009.02017.x}.

\leavevmode\hypertarget{ref-Gross2003}{}%
Gross, E. M., Feldbaum, C., and Graf, A. (2003). Epiphyte biomass and
elemental composition on submersed macrophytes in shallow eutrophic
lakes. \emph{Hydrobiologia} 506-509, 559--565.
doi:\href{https://doi.org/10.1023/B:HYDR.0000008538.68268.82}{10.1023/B:HYDR.0000008538.68268.82}.

\leavevmode\hypertarget{ref-Huerta-Cepas2017}{}%
Huerta-Cepas, J., Forslund, K., Coelho, L. P., Szklarczyk, D., Jensen,
L. J., von Mering, C., et al. (2017). Fast genome-wide functional
annotation through orthology assignment by eggNOG-mapper. \emph{Mol.
Biol. Evol.} 34, 2115--2122.
doi:\href{https://doi.org/10.1093/molbev/msx148}{10.1093/molbev/msx148}.

\leavevmode\hypertarget{ref-Huerta-Cepas2019}{}%
Huerta-Cepas, J., Szklarczyk, D., Heller, D., Hernández-Plaza, A.,
Forslund, S. K., Cook, H., et al. (2019). eggNOG 5.0: A hierarchical,
functionally and phylogenetically annotated orthology resource based on
5090 organisms and 2502 viruses. \emph{Nucleic Acids Res} 47,
D309--D314.
doi:\href{https://doi.org/10.1093/nar/gky1085}{10.1093/nar/gky1085}.

\leavevmode\hypertarget{ref-Hultman2015}{}%
Hultman, J., Waldrop, M. P., Mackelprang, R., David, M. M., McFarland,
J., Blazewicz, S. J., et al. (2015). Multi-omics of permafrost, active
layer and thermokarst bog soil microbiomes. \emph{Nature} 521, 208--212.
doi:\href{https://doi.org/10.1038/nature14238}{10.1038/nature14238}.

\leavevmode\hypertarget{ref-Hyatt2010}{}%
Hyatt, D., Chen, G.-L., LoCascio, P. F., Land, M. L., Larimer, F. W.,
and Hauser, L. J. (2010). Prodigal: Prokaryotic gene recognition and
translation initiation site identification. \emph{BMC Bioinformatics}
11, 119.
doi:\href{https://doi.org/10.1186/1471-2105-11-119}{10.1186/1471-2105-11-119}.

\leavevmode\hypertarget{ref-Jiang2015}{}%
Jiang, Y.-F., Ling, J., Dong, J.-D., Chen, B., Zhang, Y.-Y., Zhang,
Y.-Z., et al. (2015). Illumina-based analysis the microbial diversity
associated with \emph{Thalassia hemprichii} in Xincun Bay, South China
Sea. \emph{Ecotoxicology} 24, 1548--56.
doi:\href{https://doi.org/10.1007/s10646-015-1511-z}{10.1007/s10646-015-1511-z}.

\leavevmode\hypertarget{ref-Korlevic2016}{}%
Korlević, M., Šupraha, L., Ljubešić, Z., Henderiks, J., Ciglenečki, I.,
Dautović, J., et al. (2016). Bacterial diversity across a highly
stratified ecosystem: A salt-wedge Mediterranean estuary. \emph{Syst.
Appl. Microbiol.} 39, 398--408.
doi:\href{https://doi.org/10.1016/j.syapm.2016.06.006}{10.1016/j.syapm.2016.06.006}.

\leavevmode\hypertarget{ref-Kozich2013}{}%
Kozich, J. J., Westcott, S. L., Baxter, N. T., Highlander, S. K., and
Schloss, P. D. (2013). Development of a dual-index sequencing strategy
and curation pipeline for analyzing amplicon sequence data on the MiSeq
Illumina sequencing platform. \emph{Appl. Environ. Microbiol.} 79,
5112--5120.
doi:\href{https://doi.org/10.1128/AEM.01043-13}{10.1128/AEM.01043-13}.

\leavevmode\hypertarget{ref-Kuo2001}{}%
Kuo, J., and den Hartog, C. (2001). ``Seagrass taxonomy and
identification key,'' in \emph{Global Seagrass Research Methods}, eds.
F. T. Short and R. G. Coles (Amsterdam: Elsevier Science B.V.), 31--58.

\leavevmode\hypertarget{ref-Lam2016}{}%
Lam, D. W., and Lopez-Bautista, J. M. (2016). Complete chloroplast
genome for \emph{Caulerpa racemosa} (Bryopsidales, Chlorophyta) and
comparative analyses of siphonous green seaweed plastomes.
\emph{Cymbella} 2, 23--32.

\leavevmode\hypertarget{ref-Leary2014}{}%
Leary, D. H., Li, R. W., Hamdan, L. J., IV, W. J. H., Lebedev, N., Wang,
Z., et al. (2014). Integrated metagenomic and metaproteomic analyses of
marine biofilm communities. \emph{Biofouling} 30, 1211--1223.
doi:\href{https://doi.org/10.1080/08927014.2014.977267}{10.1080/08927014.2014.977267}.

\leavevmode\hypertarget{ref-Li2015}{}%
Li, D., Liu, C.-M., Luo, R., Sadakane, K., and Lam, T.-W. (2015).
MEGAHIT: An ultra-fast single-node solution for large and complex
metagenomics assembly via succinct de Bruijn graph.
\emph{Bioinformatics} 31, 1674--1676.
doi:\href{https://doi.org/10.1093/bioinformatics/btv033}{10.1093/bioinformatics/btv033}.

\leavevmode\hypertarget{ref-Li2009a}{}%
Li, H., and Durbin, R. (2009). Fast and accurate short read alignment
with Burrows--Wheeler transform. \emph{Bioinformatics} 25, 1754--1760.
doi:\href{https://doi.org/10.1093/bioinformatics/btp324}{10.1093/bioinformatics/btp324}.

\leavevmode\hypertarget{ref-Longford2007}{}%
Longford, S., Tujula, N., Crocetti, G., Holmes, A., Holmström, C.,
Kjelleberg, S., et al. (2007). Comparisons of diversity of bacterial
communities associated with three sessile marine eukaryotes.
\emph{Aquat. Microb. Ecol.} 48, 217--229.
doi:\href{https://doi.org/10.3354/ame048217}{10.3354/ame048217}.

\leavevmode\hypertarget{ref-Massana1997}{}%
Massana, R., Murray, A. E., Preston, C. M., and DeLong, E. F. (1997).
Vertical distribution and phylogenetic characterization of marine
planktonic \emph{Archaea} in the Santa Barbara Channel. \emph{Appl.
Environ. Microbiol.} 63, 50--56.

\leavevmode\hypertarget{ref-Michelou2013}{}%
Michelou, V. K., Caporaso, J. G., Knight, R., and Palumbi, S. R. (2013).
The ecology of microbial communities associated with \emph{Macrocystis
pyrifera}. \emph{PloS One} 8, e67480.
doi:\href{https://doi.org/10.1371/journal.pone.0067480}{10.1371/journal.pone.0067480}.

\leavevmode\hypertarget{ref-Morrissey2019}{}%
Morrissey, K. L., Çavas, L., Willems, A., and De Clerck, O. (2019).
Disentangling the influence of environment, host specificity and thallus
differentiation on bacterial communities in siphonous green seaweeds.
\emph{Front. Microbiol.} 10, 717.
doi:\href{https://doi.org/10.3389/fmicb.2019.00717}{10.3389/fmicb.2019.00717}.

\leavevmode\hypertarget{ref-Najdek2020}{}%
Najdek, M., Korlević, M., Paliaga, P., Markovski, M., Ivančić, I.,
Iveša, L., et al. (2020). Dynamics of environmental conditions during
the decline of a \emph{Cymodocea nodosa} meadow. \emph{Biogeosciences}
17, 3299--3315.
doi:\href{https://doi.org/10.5194/bg-17-3299-2020}{10.5194/bg-17-3299-2020}.

\leavevmode\hypertarget{ref-Neuwirth2014}{}%
Neuwirth, E. (2014). \emph{RColorBrewer: ColorBrewer palettes}.

\leavevmode\hypertarget{ref-Noges2010}{}%
Nõges, T., Luup, H., and Feldmann, T. (2010). Primary production of
aquatic macrophytes and their epiphytes in two shallow lakes (Peipsi and
Võrtsjärv) in Estonia. \emph{Aquat. Ecol.} 44, 83--92.
doi:\href{https://doi.org/10.1007/s10452-009-9249-4}{10.1007/s10452-009-9249-4}.

\leavevmode\hypertarget{ref-Parada2016}{}%
Parada, A. E., Needham, D. M., and Fuhrman, J. A. (2016). Every base
matters: Assessing small subunit rRNA primers for marine microbiomes
with mock communities, time series and global field samples.
\emph{Environ. Microbiol.} 18, 1403--1414.
doi:\href{https://doi.org/10.1111/1462-2920.13023}{10.1111/1462-2920.13023}.

\leavevmode\hypertarget{ref-Quast2013}{}%
Quast, C., Pruesse, E., Yilmaz, P., Gerken, J., Schweer, T., Yarza, P.,
et al. (2013). The SILVA ribosomal RNA gene database project: Improved
data processing and web-based tools. \emph{Nucleic Acids Res.} 41,
D590--D596.
doi:\href{https://doi.org/10.1093/nar/gks1219}{10.1093/nar/gks1219}.

\leavevmode\hypertarget{ref-RCoreTeam2019}{}%
R Core Team (2019). \emph{R: A language and environment for statistical
computing}. Vienna, Austria: R Foundation for Statistical Computing.

\leavevmode\hypertarget{ref-Reyes2001}{}%
Reyes, J., and Sansón, M. (2001). Biomass and production of the
epiphytes on the leaves of \emph{Cymodocea nodosa} in the Canary
Islands. \emph{Bot. Mar.} 44, 307--313.
doi:\href{https://doi.org/10.1515/BOT.2001.039}{10.1515/BOT.2001.039}.

\leavevmode\hypertarget{ref-Richter-Heitmann2016}{}%
Richter-Heitmann, T., Eickhorst, T., Knauth, S., Friedrich, M. W., and
Schmidt, H. (2016). Evaluation of strategies to separate root-associated
microbial communities: A crucial choice in rhizobiome research.
\emph{Front. Microbiol.} 7, 773.
doi:\href{https://doi.org/10.3389/fmicb.2016.00773}{10.3389/fmicb.2016.00773}.

\leavevmode\hypertarget{ref-Saito2019}{}%
Saito, M. A., Bertrand, E. M., Duffy, M. E., Gaylord, D. A., Held, N.
A., Hervey, W. J., et al. (2019). Progress and challenges in ocean
metaproteomics and proposed best practices for data sharing. \emph{J.
Proteome Res.} 18, 1461--1476.
doi:\href{https://doi.org/10.1021/acs.jproteome.8b00761}{10.1021/acs.jproteome.8b00761}.

\leavevmode\hypertarget{ref-Schloss2016}{}%
Schloss, P. D., Jenior, M. L., Koumpouras, C. C., Westcott, S. L., and
Highlander, S. K. (2016). Sequencing 16S rRNA gene fragments using the
PacBio SMRT DNA sequencing system. \emph{PeerJ} 4, e1869.
doi:\href{https://doi.org/10.7717/peerj.1869}{10.7717/peerj.1869}.

\leavevmode\hypertarget{ref-Schloss2009}{}%
Schloss, P. D., Westcott, S. L., Ryabin, T., Hall, J. R., Hartmann, M.,
Hollister, E. B., et al. (2009). Introducing mothur: Open-source,
platform-independent, community-supported software for describing and
comparing microbial communities. \emph{Appl. Environ. Microbiol.} 75,
7537--7541.
doi:\href{https://doi.org/10.1128/AEM.01541-09}{10.1128/AEM.01541-09}.

\leavevmode\hypertarget{ref-Staufenberger2008}{}%
Staufenberger, T., Thiel, V., Wiese, J., and Imhoff, J. F. (2008).
Phylogenetic analysis of bacteria associated with \emph{Laminaria
saccharina}. \emph{FEMS Microbiol. Ecol.} 64, 65--77.
doi:\href{https://doi.org/10.1111/j.1574-6941.2008.00445.x}{10.1111/j.1574-6941.2008.00445.x}.

\leavevmode\hypertarget{ref-Su2012}{}%
Su, C., Lei, L., Duan, Y., Zhang, K.-Q., and Yang, J. (2012).
Culture-independent methods for studying environmental microorganisms:
Methods, application, and perspective. \emph{Appl. Microbiol.
Biotechnol.} 93, 993--1003.
doi:\href{https://doi.org/10.1007/s00253-011-3800-7}{10.1007/s00253-011-3800-7}.

\leavevmode\hypertarget{ref-Uku2007}{}%
Uku, J., Björk, M., Bergman, B., and Díez, B. (2007). Characterization
and comparison of prokaryotic epiphytes associated with three East
African seagrasses. \emph{J. Phycol.} 43, 768--779.
doi:\href{https://doi.org/10.1111/j.1529-8817.2007.00371.x}{10.1111/j.1529-8817.2007.00371.x}.

\leavevmode\hypertarget{ref-Verlaque2003}{}%
Verlaque, M., Durand, C., Huisman, J. M., Boudouresque, C.-F., and Le
Parco, Y. (2003). On the identity and origin of the Mediterranean
invasive \emph{Caulerpa racemosa} (Caulerpales, Chlorophyta). \emph{Eur.
J. Phycol.} 38, 325--339.
doi:\href{https://doi.org/10.1080/09670260310001612592}{10.1080/09670260310001612592}.

\leavevmode\hypertarget{ref-Weidner1996}{}%
Weidner, S., Arnold, W., and Puhler, A. (1996). Diversity of uncultured
microorganisms associated with the seagrass \emph{Halophila stipulacea}
estimated by restriction fragment length polymorphism analysis of
PCR-amplified 16S rRNA genes. \emph{Appl. Environ. Microbiol.} 62,
766--771.

\leavevmode\hypertarget{ref-Wickham2019}{}%
Wickham, H., Averick, M., Bryan, J., Chang, W., McGowan, L. D.,
François, R., et al. (2019). Welcome to the tidyverse. \emph{J. Open
Source Softw.} 4, 1686.
doi:\href{https://doi.org/10.21105/joss.01686}{10.21105/joss.01686}.

\leavevmode\hypertarget{ref-Wilke2018}{}%
Wilke, C. O. (2018). \emph{cowplot: Streamlined plot theme and plot
annotations for 'ggplot2'}.

\leavevmode\hypertarget{ref-Williams2012}{}%
Williams, T. J., Long, E., Evans, F., DeMaere, M. Z., Lauro, F. M.,
Raftery, M. J., et al. (2012). A metaproteomic assessment of winter and
summer bacterioplankton from Antarctic Peninsula coastal surface waters.
\emph{ISME J.} 6, 1883--1900.
doi:\href{https://doi.org/10.1038/ismej.2012.28}{10.1038/ismej.2012.28}.

\leavevmode\hypertarget{ref-Wisniewski2009}{}%
Wiśniewski, J. R., Zougman, A., Nagaraj, N., and Mann, M. (2009).
Universal sample preparation method for proteome analysis. \emph{Nat.
Methods} 6, 359--362.
doi:\href{https://doi.org/10.1038/nmeth.1322}{10.1038/nmeth.1322}.

\leavevmode\hypertarget{ref-Xie2014}{}%
Xie, Y. (2014). `` knitr: A comprehensive tool for reproducible research
in R,'' in \emph{Implementing Reproducible Computational Research}, eds.
V. Stodden, F. Leisch, and R. D. Peng (New York: Chapman and Hall/CRC),
3--32.

\leavevmode\hypertarget{ref-Xie2015}{}%
Xie, Y. (2015). \emph{Dynamic Documents with R and knitr}. 2nd ed. Boca
Raton, Florida: Chapman and Hall/CRC.

\leavevmode\hypertarget{ref-Xie2019a}{}%
Xie, Y. (2019). TinyTeX: A lightweight, cross-platform, and
easy-to-maintain LaTeX distribution based on TeX Live. \emph{TUGboat}
40, 30--32.

\leavevmode\hypertarget{ref-Xie2020}{}%
Xie, Y. (2020). \emph{tinytex: Helper functions to install and maintain
'TeX Live', and compile 'LaTeX' documents}.

\leavevmode\hypertarget{ref-Xie2018}{}%
Xie, Y., Allaire, J. J., and Grolemund, G. (2018). \emph{R Markdown: The
Definitive Guide}. 1st ed. Boca Raton, Florida: Chapman and Hall/CRC.

\leavevmode\hypertarget{ref-Yilmaz2014}{}%
Yilmaz, P., Parfrey, L. W., Yarza, P., Gerken, J., Pruesse, E., Quast,
C., et al. (2014). The SILVA and "All-Species Living Tree Project (LTP)"
taxonomic frameworks. \emph{Nucleic Acids Res.} 42, D643--D648.
doi:\href{https://doi.org/10.1093/nar/gkt1209}{10.1093/nar/gkt1209}.

\leavevmode\hypertarget{ref-Zhu2019}{}%
Zhu, H. (2019). \emph{kableExtra: Construct complex table with 'kable'
and pipe syntax}.

\newpage 
\setlength\parindent{0pt}

\hypertarget{figure-legends}{%
\subsection{Figure legends}\label{figure-legends}}

\textbf{\autoref{confocal_december}.} \nameref{confocal_december}

\textbf{\autoref{confocal_june}.} \nameref{confocal_june}

\textbf{\autoref{community}.} \nameref{community}

\textbf{\autoref{chloroplast}.} \nameref{chloroplast}

\textbf{\autoref{cog}.} \nameref{cog}

\newpage

\hypertarget{figures}{%
\subsection{Figures}\label{figures}}

\begin{figure}[ht]

{\centering \includegraphics[width=0.85\linewidth]{../results/images/confocal_december}

}

\caption{Confocal microscope images of \textit{C. nodosa} and \textit{C. cylindracea} surfaces from the Bay of Saline and the Bay of Funtana (mixed and monospecific settlements) sampled on 4 December 2017 and stained with SYBR Green I. Scale bar at all images is 60 \si{\um}.\label{confocal_december}}\label{fig:unnamed-chunk-1}
\end{figure}

\newpage
\begin{figure}[ht]

{\centering \includegraphics[width=0.85\linewidth]{../results/images/confocal_june}

}

\caption{Confocal microscope images of \textit{C. nodosa} and \textit{C. cylindracea} surfaces from the Bay of Funtana (mixed and monospecific settlements) sampled on 19 June 2018 and stained with SYBR Green I. Scale bar at all images is 60 \si{\um}.\label{confocal_june}}\label{fig:unnamed-chunk-2}
\end{figure}

\newpage
\begin{figure}[ht]

{\centering \includegraphics[width=1\linewidth]{../results/figures/community_bar_plot}

}

\caption{Taxonomic classification and relative contribution of the most abundant ($\geq$ 1 \si{\percent}) bacterial and archaeal sequences from surfaces of two marine macrophytes (\textit{C. nodosa} and \textit{C. cylindracea}) sampled in the Bay of Saline and the Bay of Funtana (mixed and monospecific settlements) and in two contrasting seasons (4 December 2017 and 19 June 2018).\label{community}}\label{fig:unnamed-chunk-3}
\end{figure}

\newpage
\begin{figure}[ht]

{\centering \includegraphics[width=1\linewidth]{../results/figures/chloroplast}

}

\caption{Taxonomic classification and relative contribution of chloroplast sequences from surfaces of two marine macrophytes (\textit{C. nodosa} and \textit{C. cylindracea}) sampled in the Bay of Saline and the Bay of Funtana (mixed and monospecific settlements) and in two contrasting seasons (4 December 2017 and 19 June 2018).\label{chloroplast}}\label{fig:unnamed-chunk-4}
\end{figure}

\newpage
\begin{figure}[ht]

{\centering \includegraphics[width=1\linewidth]{../results/figures/cog}

}

\caption{Relative contribution of each COG category to the total number of annotated coding sequences (A) from metagenomes and identified proteins (B) from metaproteomes associated with surfaces of two marine macrophytes (\textit{C. nodosa} and \textit{C. cylindracea}) sampled in the Bay of Saline and the Bay of Funtana (mixed and monospecific settlements) and in two contrasting seasons (4/14 December 2017 and 19 June 2018). Total number of annotated coding sequences and identified proteins is given above the corresponding bar.\label{cog}}\label{fig:unnamed-chunk-5}
\end{figure}

\end{document}